### 2021-05-04 experiments.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{hyperref}

\title{Numerical experiments - market ecology}
\author{Aymeric Vi\'{e}}
\date{\today}

\begin{document}

\maketitle

\tableofcontents

\clearpage

\section{Preambule: baseline configuration}

Necessary condition: running GA with complete market structure
Objective: be as close as possible to Maarten's paper configuration, notably for market environment parameters.
Add Agent leverage, aggressiveness of response to be closer to Maarten
\section{Preambule 2: observables}

Necessary condition: running GA with complete market structure
To implement:
\begin{itemize}
    \item Observe market dynamics over time (easy): price, returns, assets, volatility, dividends, profits
    \item Observe ecology dynamics over time (hard): proportions? heatmap?
    Perhaps H (for max time horizon) plots with clear color codes (like jet) over time, and hoepfully we see something clear happen.
    \item Develop a GUI or something nice/updating iteself over time to observe that?
\end{itemize}

\section{Trend follower market ecology}

Necessary condition: a running GA with a single asset, market clearing, and adequate starting conditions, and the baseline configuration.

\subsection{Influence of learning vs non-learning}
The non-lerning case (replacement dynamics only) can be illustrated with:
\begin{itemize}
    \item No selection (selection rate equal to 0)
    \item No crossover (crossover rate equal to 0)
    \item No mutation (mutation rate equal to 0)
\end{itemize}
The objective is to assess the behavior of the model in this no-learning setup, establish a benchmark.

Then, we can introduce learning at different degrees, and channels. Learning requires activation of selection.

\begin{enumerate}
    \item Mutation alone, with different values
    \item Crossover alone, with different values
    \item Mutation and crossover
\end{enumerate}

How does learning affect market dynamics? Ecology dynamics?

There are many many things we can also study:
\begin{itemize}
    \item Influence of strategy space with H
    \item Time horizon of EMA fitness profit
    \item Market parameters: dividend growth rate, volatility, interest rate
    \item Alternative values for agent leverage, aggressiveness of response 
\end{itemize}

\end{document}

### 2021-05-04 experiments.tex

@@
 
-There are many many things we can also study:
+There are many many things we can also study just with trend followers:
 \begin{itemize}
@@
     \item Time horizon of EMA fitness profit
-    \item Market parameters: dividend growth rate, volatility, interest rate
+    \item Market parameters: dividend growth rate, volatility, interest rate, autocorrelation
     \item Alternative values for agent leverage, aggressiveness of response 
+    \item Population size
+    \item Tournament size
+    \item Reinvestment rate
 \end{itemize}
 
+The key area of progress will be to add multiple strategy types, this may require a lot of work on the structure. 3 separate GAs with fixed population sizes? dynamic population sizes?
+
 \end{document}

### 2021-05-05 main.py
tests to fix dividends

@@ -18,6 +18,7 @@ INITIAL_PRICE = parameters.INITIAL_PRICE
 TOURNAMENT_SIZE = parameters.TOURNAMENT_SIZE 
 INITIAL_DIVIDEND = parameters.INITIAL_DIVIDEND
 INTEREST_RATE = parameters.INTEREST_RATE
+DIVIDEND_GROWTH_RATE_G = parameters.DIVIDEND_GROWTH_RATE_G
 
 
 def main():
@@ -74,10 +75,10 @@ def main():
         @Maarten: where does the extra money from f, r, D(t) go? In the cash?
         D) I'll need to write the dividends, f, r allocation mechanism
         '''
-        market.determine_dividend_growth()
+        DIVIDEND_GROWTH_RATE = market.determine_dividend_growth(DIVIDEND_GROWTH_RATE_G)
         
         global dividend
-        dividend, random_dividend = market.draw_dividend()
+        dividend, random_dividend = market.draw_dividend(DIVIDEND_GROWTH_RATE)
         dividend_history.append(dividend)
         random_dividend_history.append(random_dividend)
         

### 2021-05-05 market.py

@@ -24,12 +24,12 @@ def truncate(number, digits) -> float:
     stepper = 10.0 ** digits
     return math.trunc(stepper * number) / stepper
 
-def determine_dividend_growth():
+def determine_dividend_growth(DIVIDEND_GROWTH_RATE_G):
     global DIVIDEND_GROWTH_RATE
     DIVIDEND_GROWTH_RATE = (1 + DIVIDEND_GROWTH_RATE_G) ** (1 / TRADING_DAYS) - 1
     return DIVIDEND_GROWTH_RATE
 
-def draw_dividend():
+def draw_dividend(DIVIDEND_GROWTH_RATE):
     
     global dividend
     global random_dividend

### 2021-05-05 parameters.py

@@ -6,7 +6,7 @@ RANDOM_SEED = random.random()
 ''' GA parmeters '''
 POPULATION_SIZE = 4
 MUTATION_RATE = 0
-MAX_GENERATIONS = 100000
+MAX_GENERATIONS = 5000
 CROSSOVER_RATE = 0
 TOURNAMENT_SIZE = 3
 

### 2021-05-05 main.py
dividends V2.0
they look fixed
at least they don't always look massively off

@@ -111,7 +111,7 @@ def main():
         # price = market_clearing_function()
         
         market.compute_ema(pop)
-        print("after ema")
+        # print("after ema")
         
         fitnessValues = list(map(ga.toolbox.evaluate, pop))
         for individual, fitnessValue in zip(pop, fitnessValues):

### 2021-05-05 market.py

@@ -26,9 +26,10 @@ def truncate(number, digits) -> float:
 
 def determine_dividend_growth(DIVIDEND_GROWTH_RATE_G):
     global DIVIDEND_GROWTH_RATE
-    DIVIDEND_GROWTH_RATE = (1 + DIVIDEND_GROWTH_RATE_G) ** (1 / TRADING_DAYS) - 1
+    DIVIDEND_GROWTH_RATE = ((1 + DIVIDEND_GROWTH_RATE_G) ** (1 / TRADING_DAYS)) - 1
     return DIVIDEND_GROWTH_RATE
 
+'''
 def draw_dividend(DIVIDEND_GROWTH_RATE):
     
     global dividend
@@ -37,17 +38,32 @@ def draw_dividend(DIVIDEND_GROWTH_RATE):
     if len(random_dividend_history) > DIVIDEND_ATC_TAU:
         random_dividend = (1 - DIVIDEND_AUTOCORRELATION ** 2) * random.normalvariate(0,1) + DIVIDEND_AUTOCORRELATION * random_dividend_history[-1 - DIVIDEND_ATC_TAU]
     dividend = abs(dividend + DIVIDEND_GROWTH_RATE * dividend + DIVIDEND_GROWTH_VOLATILITY * dividend * random_dividend)
-    print("Dividend today is " + str(dividend))
+    # print("Dividend today is " + str(dividend))
+    return dividend, random_dividend
+'''
+
+def draw_dividend(DIVIDEND_GROWTH_RATE):
+    
+    global dividend
+    global random_dividend
+    
+    random_dividend = random.normalvariate(0, 1)
+    if len(random_dividend_history) > DIVIDEND_ATC_TAU:
+        random_dividend = (1 - DIVIDEND_AUTOCORRELATION ** 2) * random_dividend + DIVIDEND_AUTOCORRELATION * random_dividend_history[len(random_dividend_history) - 1 - DIVIDEND_ATC_TAU]
+    # wiener.append(random_dividend)
+    
+    dividend = abs(dividend + DIVIDEND_GROWTH_RATE * dividend + DIVIDEND_GROWTH_VOLATILITY * dividend * random_dividend)
+    
     return dividend, random_dividend
 
         
 def wealth_earnings(pop):
-    print(INTEREST_RATE)
-    print(REINVESTMENT_RATE)
+    # print(INTEREST_RATE)
+    # print(REINVESTMENT_RATE)
     for ind in pop:
[…]

### 2021-05-05 parameters.py

@@ -6,7 +6,7 @@ RANDOM_SEED = random.random()
 ''' GA parmeters '''
 POPULATION_SIZE = 4
 MUTATION_RATE = 0
-MAX_GENERATIONS = 5000
+MAX_GENERATIONS = 100_000
 CROSSOVER_RATE = 0
 TOURNAMENT_SIZE = 3
 
@@ -25,9 +25,9 @@ INITIAL_PRICE = 1
 REINVESTMENT_RATE = 1
 INTEREST_RATE = 0.01 / TRADING_DAYS
 
-EMA_HORIZON = 20 #252
+EMA_HORIZON = 2 * TRADING_DAYS
 
-# DIVIDEND_GROWTH_RATE = 0.02 / TRADING_DAYS
+
 DIVIDEND_GROWTH_RATE_G = 0.02
 DIVIDEND_GROWTH_VOLATILITY = 0.06 / np.sqrt(TRADING_DAYS)
 DIVIDEND_AUTOCORRELATION = 0.1 

### 2021-05-05 main.py
dividend not done

@@ -40,6 +40,7 @@ def main():
     agent0_profit = []
     agent0_ema = []
     
+    dividend = INITIAL_DIVIDEND
     dividend_history.append(INITIAL_DIVIDEND)
     
     print(pop)
@@ -77,7 +78,7 @@ def main():
         '''
         DIVIDEND_GROWTH_RATE = market.determine_dividend_growth(DIVIDEND_GROWTH_RATE_G)
         
-        global dividend
+        # global dividend
         dividend, random_dividend = market.draw_dividend(DIVIDEND_GROWTH_RATE)
         dividend_history.append(dividend)
         random_dividend_history.append(random_dividend)

### 2021-05-05 experiments.py
add selection proba

@@ -4,7 +4,7 @@ import parameters
 MAX_GENERATIONS = parameters.MAX_GENERATIONS
 
 
-initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, random_dividend_history = main.main()
+initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, random_dividend_history = main.main(0)
 
 # Plot population histograms at the start and at the end
 print("--------------------------")

### 2021-05-05 main.py

@@ -1,6 +1,7 @@
 import random
 import seaborn as sns
 sns.set_theme(style="darkgrid")
+import numpy as np
 import genetic_algorithm_functions as ga
 import market as market
 import parameters
@@ -21,7 +22,7 @@ INTEREST_RATE = parameters.INTEREST_RATE
 DIVIDEND_GROWTH_RATE_G = parameters.DIVIDEND_GROWTH_RATE_G
 
 
-def main():
+def main(selection_proba):
     random.seed(RANDOM_SEED)
     
     # Create the population and the results accumulators
@@ -127,34 +128,35 @@ def main():
         
         '''
         Need to add a probability of selection (1 or 0 is enough) to be able to model a case without any learning.
-        If selection_priba is 0, offspring is just a copy of the population.
+        If selection_proba is 0, offspring is just a copy of the population.
         
         We will also be needing to add these key parameters as arguments to main
         '''
+        if selection_proba == 1:
+            offspring = ga.toolbox.select(pop, POPULATION_SIZE, TOURNAMENT_SIZE)
+           
+            print("offspring")
+            print(offspring)
+            ga.fitness_for_invalid(offspring)
+            offspring = list(map(ga.toolbox.clone, offspring))
+            
+            # Crossover
+            for child1, child2 in zip(offspring[::2], offspring[1::2]):
+                ga.toolbox.mate(child1,child2,CROSSOVER_RATE)
+                del child1.fitness.values
+                del child2.fitness.values
+            
+            # Mutation
+            for mutant in offspring:
+                ga.toolbox.mutate(mutant, MUTATION_RATE)
+                del mutant.fitness.values
         
-        offspring = ga.toolbox.select(pop, POPULATION_SIZE, TOURNAMENT_SIZE)
-        print("offspring")
-        print(offspring)
-        ga.fitness_for_invalid(offspring)
-        offspring = list(map(ga.toolbox.clone, offspring))
-        
-        # Crossover
-        for child1, child2 in zip(offspring[::2], offspring[1::2]):
-            ga.toolbox.mate(child1,child2,CROSSOVER_RATE)
-            del child1.fitness.values
-            del child2.fitness.values
-        
-        # Mutation
-        for mutant in offspring:
-            ga.toolbox.mutate(mutant, MUTATION_RATE)
-            del mutant.fitness.values
+            # Recomputing fitness
+            ga.fitness_for_invalid(offspring)
     
-        # Recomputing fitness
[…]

### 2021-05-05 parameters.py

@@ -6,7 +6,7 @@ RANDOM_SEED = random.random()
 ''' GA parmeters '''
 POPULATION_SIZE = 4
 MUTATION_RATE = 0
-MAX_GENERATIONS = 100_000
+MAX_GENERATIONS = 1_000
 CROSSOVER_RATE = 0
 TOURNAMENT_SIZE = 3
 

### 2021-05-05 main.py
structural improvements

@@ -65,73 +65,39 @@ def main(selection_proba):
         print("Generation " + str(generationCounter))
         generationCounter += 1
         
-        
-        '''
-        Here we will need to
-        A) Draw dividends
-        B) Apply dividends, interest rate and reinvestment
-        C) Update wealth sums
-        D) Hypermutation (+ update wealth?)
-        
-    
-        @Maarten: where does the extra money from f, r, D(t) go? In the cash?
-        D) I'll need to write the dividends, f, r allocation mechanism
-        '''
+        ''' A) Draw dividends '''       
         DIVIDEND_GROWTH_RATE = market.determine_dividend_growth(DIVIDEND_GROWTH_RATE_G)
-        
-        # global dividend
         dividend, random_dividend = market.draw_dividend(DIVIDEND_GROWTH_RATE)
         dividend_history.append(dividend)
         random_dividend_history.append(random_dividend)
         
+        ''' B) Apply dividends, interest rate and reinvestment, update profit '''
         market.wealth_earnings(pop)
-        market.update_wealth(pop, price)
         
+        ''' C) Update wealth sum as a function of price '''
+        market.update_wealth(pop, price) 
+        # print(pop)
         
-        print(pop)
-
-        
-        # Hypermutation
+        ''' D) Hypermutation operator '''
         global round_replacements
         round_replacements = 0
         ga.hypermutate(pop)
         # Recomputing fitness
         ga.fitness_for_invalid(pop)
         
-        
-        '''
-        E) Update trading signals
-        F) Deduce excess demand
-        G) Clear the market
-        H) Update inventories
-        I) Update wealth
-        J) Update profits
-        K) Deduce fitness as EMA
-        J) GA
-    '''
-        
-        # price = market_clearing_function()
-        
+        ''' E) Deduce fitness as EMA ''' 
         market.compute_ema(pop)
-        # print("after ema")
-        
         fitnessValues = list(map(ga.toolbox.evaluate, pop))
         for individual, fitnessValue in zip(pop, fitnessValues):
             individual.fitness.values = fitnessValue
         fitnessValues = [individual.fitness.values[0] for individual in pop]
-    
+        #turn this into a one-line function?
         print(pop)
-        print(fitnessValues)
-    
-    
+        # print(fitnessValues)
+        
+        '''  F) GA evolution of strategies with last period fitness  '''
+
         # Selection
-        
-        '''
[…]

### 2021-05-05 experiments.py
adding trading signal update mechanism

@@ -4,7 +4,7 @@ import parameters
 MAX_GENERATIONS = parameters.MAX_GENERATIONS
 
 
-initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, random_dividend_history = main.main(0)
+initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history = main.main(0)
 
 # Plot population histograms at the start and at the end
 print("--------------------------")
@@ -50,6 +50,13 @@ plt.plot(dividend_history, color='red', label='dividend')
 plt.xlabel('Generations')
 plt.ylabel('Dividend')
 plt.title('Dividends over time')
+plt.legend()
+plt.show()
+
+plt.plot(price_history, color='red', label='price')
+plt.xlabel('Generations')
+plt.ylabel('Price')
+plt.title('Prices over time')
 plt.legend()
 plt.show()
 

### 2021-05-05 main.py

@@ -35,6 +35,7 @@ def main(selection_proba):
     replacements = []
     dividend_history = []
     random_dividend_history = []
+    price_history = []
 
     
     # Temp
@@ -126,12 +127,22 @@ def main(selection_proba):
                 
         
         ''' G) Actions are now set. Update trading signals '''
-        
+        # Agent representaiton:
+#     [Theta Wealth Cash Asset Loan TradingSignal ExcessDemand     Profit     EMA profit]
+#     [ 0       1     2    3     4         5             6           7            8 ]
+
+        market.update_trading_signal(pop, price_history)
         
         ''' H) Deduce excess demand ''' 
         
         ''' I) Clear the market ''' 
         # In progress with Maarten
+        # Outputs new_price
+        
+        #temp
+        new_price = 0
+        
+        price_history.append(new_price)
         
         ''' J) Update inventories ''' 
         
@@ -156,4 +167,4 @@ def main(selection_proba):
         #         ind[1] -= 1
 
     
-    return initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, random_dividend_history
+    return initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history

### 2021-05-05 market.py

@@ -1,6 +1,7 @@
 import math
 import random
 import parameters
+import numpy as np
 
 REINVESTMENT_RATE = parameters.REINVESTMENT_RATE
 INTEREST_RATE = parameters.INTEREST_RATE
@@ -78,4 +79,17 @@ def update_wealth(pop, price):
 def compute_ema(pop):
     for ind in pop:
         ind[8] = (2 / (EMA_HORIZON + 1)) * (ind[7] - ind[8]) + ind[8]
+    return ind
+
+        # Agent representaiton:
+#     [Theta Wealth Cash Asset Loan TradingSignal ExcessDemand     Profit     EMA profit]
+#     [ 0       1     2    3     4         5             6           7            8 ]
+
+def update_trading_signal(pop, price_history):
+    for ind in pop:
+        if len(price_history) >= ind[1]:
+            ind[5] = np.log2(price_history[-1]) - np.log2(price_history[-1 - ind[1]])
+        if len(price_history) < ind[1]:
+            # The trader does not have the information to run her strategy. She waits in indifference.
+            ind[5] = 0
     return ind

### 2021-05-05 main.py
trading_signal complete
and tested

@@ -64,7 +64,7 @@ def main(selection_proba):
     while generationCounter < MAX_GENERATIONS:
         print("--------------------------")
         print("Generation " + str(generationCounter))
-        generationCounter += 1
+        
         
         ''' A) Draw dividends '''       
         DIVIDEND_GROWTH_RATE = market.determine_dividend_growth(DIVIDEND_GROWTH_RATE_G)
@@ -140,7 +140,7 @@ def main(selection_proba):
         # Outputs new_price
         
         #temp
-        new_price = 0
+        new_price = generationCounter
         
         price_history.append(new_price)
         
@@ -159,7 +159,7 @@ def main(selection_proba):
         #  Could this print results be automated? We have it twice
         
         print("- Generation {}: Max Fitness = {}, Avg Fitness = {}".format(generationCounter, maxFitness, meanFitness))
-        
+        generationCounter += 1
                 
         # Temporary function to apply some fixed cost
         # if generationCounter > 0:

### 2021-05-05 market.py

@@ -86,10 +86,19 @@ def compute_ema(pop):
 #     [ 0       1     2    3     4         5             6           7            8 ]
 
 def update_trading_signal(pop, price_history):
+    print("pop and price_history")
+    print(pop)
+    print(price_history)
     for ind in pop:
-        if len(price_history) >= ind[1]:
-            ind[5] = np.log2(price_history[-1]) - np.log2(price_history[-1 - ind[1]])
-        if len(price_history) < ind[1]:
-            # The trader does not have the information to run her strategy. She waits in indifference.
+        print(ind)
+        print(-1 - ind[0])
+        if len(price_history) > 1:
+            print(price_history[-1])
+            if len(price_history) > ind[0]:
+                ind[5] = np.log2(price_history[-1]) - np.log2(price_history[-ind[0]])
+            if len(price_history) <= ind[0]:
+                # The trader does not have the information to run her strategy. She waits in indifference.
+                ind[5] = 0
+        if len(price_history) <= 1:
             ind[5] = 0
     return ind

### 2021-05-05 parameters.py

@@ -11,7 +11,7 @@ CROSSOVER_RATE = 0
 TOURNAMENT_SIZE = 3
 
 ''' STRATEGY PARAMETERS '''
-MIN_TIME_HORIZON = 1
+MIN_TIME_HORIZON = 2
 MAX_TIME_HORIZON = 10
 
 ''' MARKET PARAMETERS '''

### 2021-05-05 main.py
addded 1.0 excess_demand computation
tested

@@ -127,13 +127,10 @@ def main(selection_proba):
                 
         
         ''' G) Actions are now set. Update trading signals '''
-        # Agent representaiton:
-#     [Theta Wealth Cash Asset Loan TradingSignal ExcessDemand     Profit     EMA profit]
-#     [ 0       1     2    3     4         5             6           7            8 ]
-
         market.update_trading_signal(pop, price_history)
         
         ''' H) Deduce excess demand ''' 
+        market.update_excess_demand(pop)
         
         ''' I) Clear the market ''' 
         # In progress with Maarten

### 2021-05-05 market.py

@@ -16,6 +16,8 @@ random_dividend = parameters.INITIAL_RANDOM_DIVIDEND
 dividend = parameters.INITIAL_DIVIDEND
 DIVIDEND_ATC_TAU = parameters.DIVIDEND_ATC_TAU
 TRADING_DAYS = parameters.TRADING_DAYS
+LAMBDA_TF = parameters.LAMBDA_TF
+STRATEGY_AGGRESSIVENESS_TF = parameters.STRATEGY_AGGRESSIVENESS_TF
 
 # Agent representaiton:
 #     [Theta Wealth Cash Asset Loan TradingSignal ExcessDemand     Profit     EMA profit]
@@ -86,14 +88,12 @@ def compute_ema(pop):
 #     [ 0       1     2    3     4         5             6           7            8 ]
 
 def update_trading_signal(pop, price_history):
-    print("pop and price_history")
-    print(pop)
-    print(price_history)
+    '''
+    Will require an update once we add different strategies
+    '''
+
     for ind in pop:
-        print(ind)
-        print(-1 - ind[0])
         if len(price_history) > 1:
-            print(price_history[-1])
             if len(price_history) > ind[0]:
                 ind[5] = np.log2(price_history[-1]) - np.log2(price_history[-ind[0]])
             if len(price_history) <= ind[0]:
@@ -101,4 +101,9 @@ def update_trading_signal(pop, price_history):
                 ind[5] = 0
         if len(price_history) <= 1:
             ind[5] = 0
+    return ind
+
+def update_excess_demand(pop):
+    for ind in pop:
+        ind[6] = ind[1] * LAMBDA_TF * (np.tanh(STRATEGY_AGGRESSIVENESS_TF * ind[5]) + 0.5)  - ind[3]
     return ind

### 2021-05-05 parameters.py

@@ -13,6 +13,8 @@ TOURNAMENT_SIZE = 3
 ''' STRATEGY PARAMETERS '''
 MIN_TIME_HORIZON = 2
 MAX_TIME_HORIZON = 10
+LAMBDA_TF = 1
+STRATEGY_AGGRESSIVENESS_TF = 1
 
 ''' MARKET PARAMETERS '''
 TRADING_DAYS = 252

### 2021-05-05 market.py
important caveat to ED

@@ -106,4 +106,6 @@ def update_trading_signal(pop, price_history):
 def update_excess_demand(pop):
     for ind in pop:
         ind[6] = ind[1] * LAMBDA_TF * (np.tanh(STRATEGY_AGGRESSIVENESS_TF * ind[5]) + 0.5)  - ind[3]
+    ''' This will need to add the division by the price '''
+    ''' This will probably have to be translated into a function '''
     return ind

### 2021-05-06 genetic_algorithm_functions.py
reworked initial wealth values

@@ -17,11 +17,11 @@ MAX_TIME_HORIZON = parameters.MAX_TIME_HORIZON
 MUTATION_RATE = parameters.MUTATION_RATE
 MAX_GENERATIONS = parameters.MAX_GENERATIONS
 CROSSOVER_RATE = parameters.CROSSOVER_RATE
-MIN_WEALTH = parameters.MIN_WEALTH
-MAX_WEALTH = parameters.MAX_WEALTH
 MIN_TIME_HORIZON = parameters.MIN_TIME_HORIZON
 INITIAL_PRICE = parameters.INITIAL_PRICE
 TOURNAMENT_SIZE = parameters.TOURNAMENT_SIZE 
+INITIAL_ASSETS = parameters.INITIAL_ASSETS
+INITIAL_CASH = parameters.INITIAL_CASH
 
 # =============================================================================
 # Setup the evolutionary operators
@@ -40,8 +40,8 @@ creator.create("individual", list, typecode = 'd', fitness=creator.fitness_strat
 # Create the individual list 
 toolbox.register("generate_strategy", random.randint, MIN_TIME_HORIZON, MAX_TIME_HORIZON)
 toolbox.register("generate_wealth", random.randint, 0, 0)
-toolbox.register("generate_cash", random.randint, 5, 10)
-toolbox.register("generate_asset", random.randint, 5, 5)
+toolbox.register("generate_cash", random.randint, INITIAL_CASH, INITIAL_CASH)
+toolbox.register("generate_asset", random.randint, INITIAL_ASSETS, INITIAL_ASSETS)
 toolbox.register("generate_loan", random.randint, 0, 0)
 toolbox.register("generate_trading_signal", random.randint, 0, 0)
 toolbox.register("generate_excess_demand", random.randint, 0, 0)

### 2021-05-06 main.py

@@ -12,8 +12,6 @@ MAX_TIME_HORIZON = parameters.MAX_TIME_HORIZON
 MUTATION_RATE = parameters.MUTATION_RATE
 MAX_GENERATIONS = parameters.MAX_GENERATIONS
 CROSSOVER_RATE = parameters.CROSSOVER_RATE
-MIN_WEALTH = parameters.MIN_WEALTH
-MAX_WEALTH = parameters.MAX_WEALTH
 MIN_TIME_HORIZON = parameters.MIN_TIME_HORIZON
 INITIAL_PRICE = parameters.INITIAL_PRICE
 TOURNAMENT_SIZE = parameters.TOURNAMENT_SIZE 

### 2021-05-06 parameters.py

@@ -19,21 +19,20 @@ STRATEGY_AGGRESSIVENESS_TF = 1
 ''' MARKET PARAMETERS '''
 TRADING_DAYS = 252
 
-MIN_WEALTH = 10
-MAX_WEALTH = 10
-
-INITIAL_PRICE = 1
-
+INITIAL_PRICE = 100
+INITIAL_CASH = 50_000_000
+INITIAL_ASSETS = 500_000
+ # = Wealth of 100M; 100M is the threshold for US asset managers
+ 
 REINVESTMENT_RATE = 1
 INTEREST_RATE = 0.01 / TRADING_DAYS
-
 EMA_HORIZON = 2 * TRADING_DAYS
 
 
 DIVIDEND_GROWTH_RATE_G = 0.02
 DIVIDEND_GROWTH_VOLATILITY = 0.06 / np.sqrt(TRADING_DAYS)
 DIVIDEND_AUTOCORRELATION = 0.1 
-INITIAL_DIVIDEND = 0.01
+INITIAL_DIVIDEND = 100
 INITIAL_RANDOM_DIVIDEND = 0
 DIVIDEND_ATC_TAU = 3
 

### 2021-05-06 experiments.py
visual update for experiments

@@ -19,10 +19,17 @@ print("--------------------------")
 # Plot the fitness evolution over time
 plt.plot(maxFitnessValues, color='red', label='Maximum fitness')
 plt.plot(meanFitnessValues, color='green', label = 'Average fitness')
-plt.plot(replacements, color='gray', label = 'Hypermutations')
 plt.xlabel('Generations')
 plt.ylabel('Max / Average Fitness')
 plt.title('Max and Average Fitness over Generations')
+plt.xlim(0,MAX_GENERATIONS+1)
+plt.legend()
+plt.show()
+
+plt.plot(replacements, color='gray', label = 'Hypermutations')
+plt.xlabel('Generations')
+plt.ylabel('Hypermutations')
+plt.title('Hypermutations over time')
 plt.xlim(0,MAX_GENERATIONS+1)
 plt.legend()
 plt.show()
@@ -53,7 +60,7 @@ plt.title('Dividends over time')
 plt.legend()
 plt.show()
 
-plt.plot(price_history, color='red', label='price')
+plt.plot(price_history, color='black', label='price')
 plt.xlabel('Generations')
 plt.ylabel('Price')
 plt.title('Prices over time')

### 2021-05-06 market_clearing.py
solving model one by active thread issue

@@ -59,6 +59,7 @@ Short selling
 from esl.economics.markets.walras import excess_demand_model, differentiable_order_message
 model = excess_demand_model({p: initial_quote, p2: initial_quote})
 
+
 market_agent = esl.simulation.identity([0])
 trader_agent = esl.simulation.identity([1])
 
@@ -129,4 +130,8 @@ for k, v in multipliers.items():
 
 
 print("We can access the numeric value of the price with float(new_price) = " + str(float(new_price)))
+
+''' IMPORTANT: We may have to del model afterwards, or put it inside a function, 
+so that we don't have the error attempt to activate an adept when one is already active in this thread'''
+del model
 

### 2021-05-06 experiments.py
adding arguments to main for learning/no learning experiments

@@ -3,8 +3,12 @@ import main
 import parameters
 MAX_GENERATIONS = parameters.MAX_GENERATIONS
 
+# Optional because they are arguments to main
+CROSSOVER_RATE = parameters.CROSSOVER_RATE
+MUTATION_RATE = parameters.MUTATION_RATE 
 
-initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history = main.main(0)
+
+initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history = main.main(1, CROSSOVER_RATE, MUTATION_RATE)
 
 # Plot population histograms at the start and at the end
 print("--------------------------")

### 2021-05-06 main.py

@@ -9,9 +9,9 @@ import parameters
 RANDOM_SEED = parameters.RANDOM_SEED
 POPULATION_SIZE = parameters.POPULATION_SIZE
 MAX_TIME_HORIZON = parameters.MAX_TIME_HORIZON
-MUTATION_RATE = parameters.MUTATION_RATE
+# MUTATION_RATE = parameters.MUTATION_RATE
 MAX_GENERATIONS = parameters.MAX_GENERATIONS
-CROSSOVER_RATE = parameters.CROSSOVER_RATE
+# CROSSOVER_RATE = parameters.CROSSOVER_RATE
 MIN_TIME_HORIZON = parameters.MIN_TIME_HORIZON
 INITIAL_PRICE = parameters.INITIAL_PRICE
 TOURNAMENT_SIZE = parameters.TOURNAMENT_SIZE 
@@ -20,7 +20,7 @@ INTEREST_RATE = parameters.INTEREST_RATE
 DIVIDEND_GROWTH_RATE_G = parameters.DIVIDEND_GROWTH_RATE_G
 
 
-def main(selection_proba):
+def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     random.seed(RANDOM_SEED)
     
     # Create the population and the results accumulators

### 2021-05-06 experiments.tex
Updates from Overleaf

@@
 \usepackage{hyperref}
+\newcommand{\code}{\texttt}
+\usepackage[
+backend=biber,
+style=alphabetic,
+sorting=ynt
+]{biblatex}
+\addbibresource{bibliography.bib}
 
@@
 
-\section{Preambule: baseline configuration}
+\section{Current Model description}
+
+\subsection{Initialisation}
+
+We first create a population of $n$ trend-following agents, at time $t=0$.
+Each agent $i$ is described by its strategy $\theta_i(0)$, initial cash $C_i(0)$, initial assets $S_i(0)$.
+The $\theta$ parameter describes the time horizon on which the trend following strategy is based (Equation \ref{trend_following_equation}).
+Initial strategies are drawn uniformly in $(1,H]$, where $H$ is the maximum possible time horizon.
+\par
+Values for initial endowments of the agents meet the $\$100M$ Securities and Exchange Commission threshold for institutional investment managers, that are equally split in cash and asset shares.
+
+\subsection{Market structure}
+
+In this market, agents are choosing between a bond with a fixed continuously paid interest rate $r = 1\%$ annually, and an asset which pays dividends $D(t)$ at each step $t$, modelled as an autocorrelated geometric Brownian Motion.
+
+\begin{equation}
+    \label{dividend_equation}
+    \begin{array}{l}
+{D}(t)={D}(t)+g {D}(t-1)+\sigma {D}(t-1) {U}(t), \\
+{U}(t)=\omega U(t-2)+\left(1-\omega^{2}\right){Z}(t)
+\end{array}
+\end{equation}
+
+Where $g$ is the dividend growth rate, $\sigma$ its variance, $\omega$ is the autocorrelation parameter of the process, and $Z$ is a standard Wiener process.
+Based on estimates from Lebaron \cite{lebaron2001empirical} and Scholl et al.
+\cite{scholl2020market}, we take $g=2\%$ annually and $\sigma = 6\%$ annually.
+\par
+The market is initialised with an initial price of $100$ that matches the fundamental value of the asset, and initial dividends of $100$.
+
+\subsection{Strategy and market behavior}
+% here we write what the agents do, trading signal (write an example with a time series and a few different agents?
+that would be very nice.
+Inspire from David's figure 1?), excess demand, market clearing
+% mention the next strategy extensions
+% mention hypermutation & replacement
+
+\begin{equation}
+    \label{trend_following_equation}
+
+\end{equation}
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{c|c|c}
+    Mathematical notation & Meaning & Value \\
+    \hline
+        $n$ & Number of agents & 100 \\
+        $t$ & Time index & $[1,t_{\text{max}}]$\\
+        $\theta_i(t)$ & Agent strategy at time $t$ & $(1, H]$\\
+        $C_i(0)$ & Initial cash & 50,000,000 \\
+        $S_i(0)$ & Initial asset shares & 500,000 \\
+        $D(t)$ & Dividend per asset share & See equation \ref{dividend_equation} \\
+        $g$ & Dividend growth rate & $0.02$ annually \\
+        $\sigma$ & Dividend variance & $0.06$ annually \\
+        $\omega$ & Dividend autocorrelation & $0.1$ \\
+    \end{tabular}
+    \caption{Parameters and notations}
+    \label{parameters_correspondence}
+\end{table}
+
+\subsection{Evolution of strategies}
+% here we describe fitness as EMA, the GA operators, what the GA imitates (learning, imitation).
+
+\section{Numerical experiment: identifying the impact of evolution in market dynamics}
+% comparison between evolution and non evolution
+% need a clear plan, not necessarily all finished, but at least for what's unclear (how many repetitions? How many agents?), some ideas to clear this out.
+
+\section{Preamble: baseline configuration}
 
@@
 Add Agent leverage, aggressiveness of response to be closer to Maarten
-\section{Preambule 2: observables}
+\section{Preamble 2: observables}
 
@@
     \item Observe market dynamics over time (easy): price, returns, assets, volatility, dividends, profits
-    \item Observe ecology dynamics over time (hard): proportions? heatmap?
-    Perhaps H (for max time horizon) plots with clear color codes (like jet) over time, and hoepfully we see something clear happen.
-    \item Develop a GUI or something nice/updating iteself over time to observe that?
+    \item Observe ecology dynamics over time (hard): proportions? heatmap? Add entropy?
+    Perhaps H (for max time horizon) plots with clear color codes (like jet) over time, and hopefully we see something clear happen.
+    \item Develop a GUI or something nice/updating itself over time to observe that?
 \end{itemize}
@@
 
+\printbibliography
+
 \end{document}

### 2021-05-10 experiments.tex
Updates from Overleaf

@@
 
-\title{Numerical experiments - market ecology}
+\title{Evolutionary market ecologies \\ Model \& Numerical experiments}
 \author{Aymeric Vi\'{e}}
@@
 
-\section{Current Model description}
+\section{Trend following Model description}
 
@@
 \subsection{Strategy and market behavior}
+At each time period, trend followers look at past prices to decide the content of their next portfolio, choosing between cash and an asset.
+The share of their wealth that will be devoted to asset shares is proportional to the agent trading signal $\phi(t)$.
+Here, the agent compares the logarithm of the prices at last period, and at another period $\theta$.
+This $\theta$ is the time horizon of this agent's trend following strategy, and is subject to evolution.
+We consider $\theta \in (1, H]$ with $H$ the maximum allowed time horizon.
+In \cite{scholl2020market}, trend followers all had $\theta = 2$; this generalisation allows us to look at a larger space of strategies.
 % here we write what the agents do, trading signal (write an example with a time series and a few different agents?
@@
     \label{trend_following_equation}
+    \phi_i(t) = \log_2 p(t-1) - \log_2 p(t-\theta_i)
+\end{equation}
 
+Once the trading signals have been updated, the agents express their excess demand for the asset, as shown in Equation \ref{excess_demand_equation}.
+
+\begin{equation}
+\label{excess_demand_equation}
+   E_i(t) = \frac{W_i(t) {\lambda_i}}{p(t)} \left( \tanh{({c} \phi_i(t)}) + \frac{1}{2} \right) - S_i(t)
 \end{equation}
+
+The use of $\tanh$ ensures that the function is bounded and differentiable, for the use of the market clearing algorithm.
+$c$ denotes the aggressiveness of the agent strategic response, i.e.
+how much its demand is sensitive to the trading signal.
+$\lambda$ is the agent leverage.
+The term $\frac{1}{2}$ ensures that when the agent is indifferent to assets (i.e.
+when $\phi(t) = 0$), then the agent equally splits its wealth into cash and asset shares.
+
+\subsection{Market clearing}
+
+Once all agents have formulated their excess demand functions as a function of the unknown current price $p(t)$, we use Walrasian market clearing to identify the clearing price, as described in \cite{scholl2020market}.
+In short, excess demands of all agents $a$ in the population $ \mathcal{A}$ are submitted to a price maker agent who attempts to find the root of the aggregate demand, i.e.
+the price such that the aggregate excess demand is equal to 0.
+
+\begin{equation}
+\label{market_clearing_root}
+\sum_{a \in \mathcal{A}} E_{a}(p)=0
+\end{equation}
+
+If such root does not exist, the price maker instead minimises the mismatch between supply and demand, under the constraint that $p$ is finite.
+
+\begin{equation}
+\label{market_clearing_mismatch}
+\min_p\left(\sum_{a \in, \mathcal{A}} E_{a}(p)\right)^{2}
+\end{equation}
+
+This process allows to pin down the clearing price $p(t)$.
+We can thus compute the realised demands of the agents, update their asset shares ownership, wealth and profit.
+The agents receive dividends $D(t)$ in proportion of their asset shares, receive interest $r$ on their cash.
+Finally, investment inflows and outflows are modelled by the reinvestment parameter $f$, which is the multiplier of the agent profit, i.e.
+$W(t+1) - W(t)$.
+For $f > 1$, profits cause an inflow in capital.
+When $f < 1$, profits cause investors to withdraw funds from the agent-fund; finally, when $f=1$, no particular reinvestment dynamic occurs.
 
@@
         $n$ & Number of agents & 100 \\
-        $t$ & Time index & $[1,t_{\text{max}}]$\\
+        $t$ & Time index & $[1,T]$\\
+        $T$ & Maximum time index & $[1000;100,000]$\\
         $\theta_i(t)$ & Agent strategy at time $t$ & $(1, H]$\\
@@
         $\omega$ & Dividend autocorrelation & $0.1$ \\
+        $c$ & Strategy aggressiveness & $[1,10]$ \\
+        $\lambda$ & Agent leverage & $[1,10]$ \\
+        $f$ & Reinvestment rate & $[0,2]$ \\
+        $\tau$ & Tournament size & $(1,n]$ \\
+        $\nu$ & EMA length & $[1,2520]$ \\
+        $\rho$ & Crossover rate & $[0,1]$ \\
+        $\mu$ & Mutation rate & $[0,1]$ \\
+        $\varsigma$ & Selection rate & $\{0,1\}$
     \end{tabular}
@@
 \end{table}
+% some here are not parameters, and will have to be removed
 
 \subsection{Evolution of strategies}
-% here we describe fitness as EMA, the GA operators, what the GA imitates (learning, imitation).
+The main contribution of this model is to emphasise the evolution of strategies.
+This models funds, managers, who look at the profits and performance of other funds, and tend to imperfectly imitate stronger financial strategies.
+Each fund can also independently choose to marginally change its strategy, in an exploratory behavior.\\
+% some refs needed here to support that this behavior exists and is well documented
+
+First, we implement an \textit{hypermutation} operator in which insolvent agents are removed from the populations.
+Agents with wealth $W \leq 0$ disappear from the market, and are replaced by new individuals with random strategies and usual initial endowments.
+% they may have to be initialised only with cash because otherwise we are creating new asset shares ex nihilo.
+
+The \textit{fitness} variable looked at for imitation is the exponential moving average of profits $\eta(t)$. For an agent $i$ with current profit $\pi_i(t)$:
+
+\begin{equation}
+    \eta_i(t) =  \frac{2}{ \nu + 1} \left(\pi_i(t) -  \eta_i(t-1)\right) +  \eta_i(t-1)
+\end{equation}
+
+Where $\nu$ represents the length of the EMA, i.e.
+the number of day considered.
+Typical examples of EMAs are 10-day, 50-day or 200-day moving averages.
+Here, $\nu$ encodes a $\nu$-day EMA for profits.
+$2$ is chosen as an arbitrary smoothing factor.\\
+
+In this model, this imitation and exploration behavior are modeled using a genetic algorithm.
+We first operate a modified tournament \textit{selection} with probability $\varsigma$, the step in which each fund selects a few funds it will compare itself to.
+The number of selected funds in this tournament is equal to the tournament size $\tau$, set by default to 3.
+The $\tau$ randomly drawn funds are compared, and the one with the highest fitness is selected.
+The closer $\tau$ is to the population size $n$, the more reliable the tournament selection outputs the best performing strategy.
+For each agent $i$ in the population, a tournament selection is operated, and this agent $i$ will imitate the strategy $\theta$ of the highest fitness individual in this tournament.\\
+
+Once imitation by tournament selection is operated, we allow strategies to recombine together and explore alternative regions of the strategy space with probability $\rho$.
+This is the \textit{crossover} operator.
+In the population, pairs of individuals are gathered, such that each individual is the member of a single pair.
+For a pair with individuals $A$ and $B$ with respective strategies $^\theta_A$ and $\theta_B$, the resulting individuals $A'$ and $B'$ from the crossover operator are characterised by strategies:
+
+\begin{equation}
+    \theta_A' \sim \mathcal{U}([\min(\theta_A; \theta_B), \max(\theta_A; \theta_B)] 
+\end{equation}
+\begin{equation}
+    \theta_B' \sim \mathcal{U}([\min(\theta_A; \theta_B), \max(\theta_A; \theta_B)] 
+\end{equation}
+
+In other words, the resulting strategies from the crossover of $A$ and $B$ are uniformly drawn in the space of strategies bounded by the strategies of $A$ and $B$.
+If $A$ and $B$ had strategies $2$ and $4$, the possible integer strategies resulting from crossover would be $2, 3$ and $4$.
+For instance, $5$ would not be possible, as it lays outside of the space of strategies covered by the pair ($A,B$).\\
+
+Finally, marginal exploration behavior is allowed by the \textit{mutation} operator. With probability $\mu$,
+
+% need a table to summarise the operators, their effects and parameters involved
+% add pseudocode?
+
+\subsection{Extending to multiple strategy types \& assets}
+% with value investors and noise traders
+% and mutliple assets
 

### 2021-06-07 market_clearing.py
reorganising repo

@@ -1,47 +1,0 @@
-
-
-def aggregate_ed(values):
-    # return (3 * x - 10) for x in values
-
-    return sum([3 * x - 10 for x in values])
-price = 1
-
-'''
-Here is a simple optimisation solver for the market clering algorithm. 
-We minimise the squared of aggregate excess demand, under the constraint of
-a Limit-Up-Limit-Down (LU-LD) circuit breaker, which limits prices to 
-[LD*p(t-1), LU*p(t-1)], with LD=1/2 and LU = 2
-'''
-
-''' First define squared aggregate demand '''
-def squared_agg_ed(x):
-    result = aggregate_ed(x) ** 2
-    return result
-
-limit_down = price * 0.5
-limit_up = price * 2.0
-
-import leap_simple_modif as leapm
-
-best_genome = leapm.ea_solve_verbose_false(squared_agg_ed,
-         bounds=[(limit_down, limit_up)], generations = 1000, pop_size = 50,
-         mutation_std=0.1, hard_bounds = True)
-
-# best_genome = ea_solve(squared_agg_ed,
-#          bounds=[(limit_down, limit_up)], generations = 1000, pop_size = 50,
-#          mutation_std=0.1)
-print(best_genome)
-print(best_genome[0])
-
-
-
-# best_genome = leapm.ea_solve_verbose_false(squared_agg_ed,
-#          bounds=[(0.5, 2)], generations = 1000, pop_size = 50,
-#          mutation_std=0.1)
-
-
-
-from leap_ec.simple import ea_solve
-def f(x):
-    return sum(x)**2
-ea_solve(f, bounds=[(4, 5.12) for _ in range(1)], maximize=False, hard_bounds = True)

### 2021-06-07 market_clearing_leap.py

@@ -1,4 +1,9 @@
-''' from leap_ec.simple import ea_solve '''
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jun  7 08:38:08 2021
+
+@author: aymer
+"""
 
 """
     Provides a very high-level convenience function for a very general EA,
@@ -19,7 +24,7 @@ from leap_ec.real_rep.initializers import create_real_vector
 ##############################
 # Function ea_solve()
 ##############################
-def ea_solve_verbose_false(function, bounds, generations=100, pop_size=2,
+def ea_solve(function, bounds, generations=100, pop_size=2,
              mutation_std=1.0, maximize=False, viz=False, viz_ylim=(0, 1),
              hard_bounds=True):
     """Provides a simple, top-level interfact that optimizes a real-valued
@@ -35,7 +40,30 @@ def ea_solve_verbose_false(function, bounds, generations=100, pop_size=2,
     :param bool viz: whether to display a live best-of-generation plot
     :param bool hard_bounds: if True, bounds are enforced at all times during
         evolution; otherwise they are only used to initialize the population.
-"""
+    :param (float, float) viz_ylim: initial bounds to use of the plots
+        vertical axis
+    The basic call includes instrumentation that prints the best-so-far fitness
+    value of each generation to stdout:
+    >>> from leap_ec.simple import ea_solve
+    >>> ea_solve(sum, bounds=[(0, 1)]*5) # doctest:+ELLIPSIS
+    generation, bsf
+    0, ...
+    1, ...
+    ...
+    100, ...
+    [..., ..., ..., ..., ...]
+    When `viz=True`, a live BSF plot will also display:
+    >>> ea_solve(sum, bounds=[(0, 1)]*5, viz=True) # doctest:+ELLIPSIS
+    generation, bsf
+    0, ...
+    1, ...
+    ...
+    100, ...
+    [..., ..., ..., ..., ...]
+    .. plot::
+        from leap_ec.simple import ea_solve
+        ea_solve(sum, bounds=[(0, 1)]*5, viz=True)
+    """
 
     if hard_bounds:
         mutation_op = mutate_gaussian(std=mutation_std, hard_bounds=bounds,
@@ -53,9 +81,9 @@ def ea_solve_verbose_false(function, bounds, generations=100, pop_size=2,
         ops.pool(size=pop_size)
     ]
 
-    # if viz:
[…]

### 2021-06-07 testing_&_archiv/leap_issue.py

@@ -1,0 +1,110 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jun  7 08:38:08 2021
+
+@author: aymer
+"""
+
+"""
+    Provides a very high-level convenience function for a very general EA,
+    ea_solve().
+"""
+from matplotlib import pyplot as plt
+
+from leap_ec import ops, probe
+from leap_ec.algorithm import generational_ea
+from leap_ec.real_rep.ops import mutate_gaussian
+from leap_ec.representation import Representation
+from leap_ec.individual import Individual
+from leap_ec.problem import FunctionProblem
+from leap_ec.decoder import IdentityDecoder
+from leap_ec.real_rep.initializers import create_real_vector
+
+
+##############################
+# Function ea_solve()
+##############################
+def ea_solve(function, bounds, generations=100, pop_size=2,
+             mutation_std=1.0, maximize=False, viz=False, viz_ylim=(0, 1),
+             hard_bounds=True):
+    """Provides a simple, top-level interfact that optimizes a real-valued
+    function using a simple generational EA.
+    :param function: the function to optimize; should take lists of real
+        numbers as input and return a float fitness value
+    :param [(float, float)] bounds: a list of (min, max) bounds to define the
+        search space
+    :param int generations: the number of generations to run for
+    :param int pop_size: the population size
+    :param float mutation_std: the width of the mutation distribution
+    :param bool maximize: whether to maximize the function (else minimize)
+    :param bool viz: whether to display a live best-of-generation plot
+    :param bool hard_bounds: if True, bounds are enforced at all times during
+        evolution; otherwise they are only used to initialize the population.
+    :param (float, float) viz_ylim: initial bounds to use of the plots
+        vertical axis
+    The basic call includes instrumentation that prints the best-so-far fitness
+    value of each generation to stdout:
+    >>> from leap_ec.simple import ea_solve
+    >>> ea_solve(sum, bounds=[(0, 1)]*5) # doctest:+ELLIPSIS
+    generation, bsf
+    0, ...
+    1, ...
+    ...
+    100, ...
+    [..., ..., ..., ..., ...]
+    When `viz=True`, a live BSF plot will also display:
+    >>> ea_solve(sum, bounds=[(0, 1)]*5, viz=True) # doctest:+ELLIPSIS
+    generation, bsf
+    0, ...
+    1, ...
+    ...
+    100, ...
[…]

### 2021-06-07 testing_&_archiv/leap_simple_modif.py

@@ -1,0 +1,82 @@
+''' from leap_ec.simple import ea_solve '''
+
+"""
+    Provides a very high-level convenience function for a very general EA,
+    ea_solve().
+"""
+from matplotlib import pyplot as plt
+
+from leap_ec import ops, probe
+from leap_ec.algorithm import generational_ea
+from leap_ec.real_rep.ops import mutate_gaussian
+from leap_ec.representation import Representation
+from leap_ec.individual import Individual
+from leap_ec.problem import FunctionProblem
+from leap_ec.decoder import IdentityDecoder
+from leap_ec.real_rep.initializers import create_real_vector
+
+
+##############################
+# Function ea_solve()
+##############################
+def ea_solve_verbose_false(function, bounds, generations=100, pop_size=2,
+             mutation_std=1.0, maximize=False, viz=False, viz_ylim=(0, 1),
+             hard_bounds=True):
+    """Provides a simple, top-level interfact that optimizes a real-valued
+    function using a simple generational EA.
+    :param function: the function to optimize; should take lists of real
+        numbers as input and return a float fitness value
+    :param [(float, float)] bounds: a list of (min, max) bounds to define the
+        search space
+    :param int generations: the number of generations to run for
+    :param int pop_size: the population size
+    :param float mutation_std: the width of the mutation distribution
+    :param bool maximize: whether to maximize the function (else minimize)
+    :param bool viz: whether to display a live best-of-generation plot
+    :param bool hard_bounds: if True, bounds are enforced at all times during
+        evolution; otherwise they are only used to initialize the population.
+"""
+
+    if hard_bounds:
+        mutation_op = mutate_gaussian(std=mutation_std, hard_bounds=bounds,
+                        expected_num_mutations='isotropic')
+    else:
+        mutation_op = mutate_gaussian(std=mutation_std,
+                        expected_num_mutations='isotropic')
+
+    pipeline = [
+        ops.tournament_selection,
+        ops.clone,
+        mutation_op,
+        ops.uniform_crossover(p_swap=0.4),
+        ops.evaluate,
+        ops.pool(size=pop_size)
+    ]
+
+    # if viz:
+    #     plot_probe = probe.FitnessPlotProbe(ylim=viz_ylim, ax=plt.gca())
+    #     pipeline.append(plot_probe)
+
+    ea = generational_ea(generations=generations, pop_size=pop_size,
+                         problem=FunctionProblem(function, maximize),
+
[…]

### 2021-06-07 testing_&_archiv/market_clearing.py

@@ -1,0 +1,48 @@
+
+
+def aggregate_ed(values):
+    # return (3 * x - 10) for x in values
+
+    return sum([3 * x - 10 for x in values])
+price = 1
+
+# '''
+# Here is a simple optimisation solver for the market clering algorithm. 
+# We minimise the squared of aggregate excess demand, under the constraint of
+# a Limit-Up-Limit-Down (LU-LD) circuit breaker, which limits prices to 
+# [LD*p(t-1), LU*p(t-1)], with LD=1/2 and LU = 2
+# '''
+
+# ''' First define squared aggregate demand '''
+# def squared_agg_ed(x):
+#     result = aggregate_ed(x) ** 2
+#     return result
+
+# limit_down = price * 0.5
+# limit_up = price * 2.0
+
+# import leap_simple_modif as leapm
+
+# best_genome = leapm.ea_solve_verbose_false(squared_agg_ed,
+#          bounds=[(limit_down, limit_up)], generations = 1000, pop_size = 50,
+#          mutation_std=0.1, hard_bounds = True)
+
+# # best_genome = ea_solve(squared_agg_ed,
+# #          bounds=[(limit_down, limit_up)], generations = 1000, pop_size = 50,
+# #          mutation_std=0.1)
+# print(best_genome)
+# print(best_genome[0])
+
+
+
+# # best_genome = leapm.ea_solve_verbose_false(squared_agg_ed,
+# #          bounds=[(0.5, 2)], generations = 1000, pop_size = 50,
+# #          mutation_std=0.1)
+
+
+
+# from leap_ec.simple import ea_solve
+# def f(x):
+#     return sum(x)**2
+# ea_solve(f, bounds=[(4, 5.12) for _ in range(1)], maximize=False, hard_bounds = True)
+

### 2021-06-07 testing_&_archiv/mc_test2.py

@@ -1,0 +1,84 @@
+import esl
+from esl.law import property
+from esl.simulation import identity
+from esl.economics import price
+from esl.economics import currencies
+from esl.economics.markets import quote
+from esl.economics.markets.walras import excess_demand_model, differentiable_order_message
+
+def solve(excess_demand_functions: list):
+    i  = identity([0, 1])
+    p  = property(i)
+    initial_price = price(int(1.23 * currencies.USD.denominator), currencies.USD)
+    initial_quote = quote(initial_price)
+    market_agent = esl.simulation.identity([0])
+    excess_demand_functions = []
+    model = excess_demand_model({p: initial_quote})
+
+    # for i, ed in enumerate(my_excess_demand_functions):
+
+    #     order = ed(esl.simulation.identity([i+1]), market_agent, 0, 0)
+
+    #     excess_demand_functions.append(order) 
+
+    model.excess_demand_functions = excess_demand_functions
+    multipliers = model.compute_clearing_quotes() 
+
+    prices = []
+
+    for k, v in multipliers.items():
+
+        prices.append(price(round(float(initial_price) * v * currencies.USD.denominator), currencies.USD))
+
+
+    del model
+    return prices
+
+def aggregate_ed1(x):
+    return 3 * x - 10
+
+def aggregate_ed2(x):
+    return 5 * x - 7
+
+aggregate_ed = [aggregate_ed1, aggregate_ed2]
+
+prices = solve(aggregate_ed)
+print(prices)
+
+import esl
+from esl.law import property
+from esl.simulation import identity
+from esl.economics import price
+from esl.economics import currencies
+from esl.economics.markets import quote
+from esl.economics.markets.walras import excess_demand_model, differentiable_order_message
+def solve(my_excess_demand_functions: list):
+    
+     i  = identity([0, 1])
+     p  = property(i)
+     initial_price = price(int(1.23 * currencies.USD.denominator), currencies.USD)
+     initial_quote = quote(initial_price)
+     model = excess_demand_model({p: initial_quote})
+    
+     market_agent = esl.simulation.identity([0])
+     excess_demand_functions = []
+     for i, ed in enumerate(my_excess_demand_functions):
+         order = ed(esl.simulation.identity([i+1]), market_agent, 0, 0)
+         excess_demand_functions.append(order) 
+     model.excess_demand_functions = excess_demand_functions
+     multipliers = model.compute_clearing_quotes() 
[…]

### 2021-06-07 testing_&_archiv/mc_tests.py

@@ -1,3 +1,28 @@
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
 ''' https://test.pypi.org/project/eslpy/ '''
 ''' We want it to be 0.0.132 '''
 

### 2021-06-07 main.py
correction asset shares after clearing

@@ -47,7 +47,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     print("Initial population")
     print(('{}\n'*len(pop)).format(*pop))
     print("Agent representation")
-    print("[Theta Wealth Cash Asset Loan TradingSignal ExcessDemand     Profit     EMA profit]")
+    print("[Theta Wealth Cash Asset Loan TradingSignal RawExcessDemand     Profit     EMA profit]")
     print("[ 0       1     2    3     4         5             6           7            8 ]")
     
     fitnessValues = list(map(ga.toolbox.evaluate, pop))
@@ -77,7 +77,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         random_dividend_history.append(random_dividend)
         
         ''' B) Apply dividends, interest rate and reinvestment, update profit '''
-        market.wealth_earnings(pop, dividend)
+        # market.wealth_earnings(pop, dividend)
         
         ''' C) Update wealth sum as a function of price '''
         market.update_wealth(pop, price) 

### 2021-06-07 market.py

@@ -117,7 +117,7 @@ def order_excess_demand(pop):
     list_excess_demand_func = []
     for ind in pop:
         def ed(price):
-            return ind[6] / price
+            return ind[6] / price - ind[3]
         list_excess_demand_func.append(ed)
     return list_excess_demand_func
 
@@ -138,6 +138,6 @@ def compute_aggregate_excess_demand(pop):
 
 def assign_assets (pop, price):
     for ind in pop:
-        ind[3] += ind[6] / price - ind[3]
+        ind[3] = ind[6] / price + ind[3]
     return ind
 

### 2021-06-07 main.py
cash update

@@ -152,7 +152,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         price_history.append(price)
         
         ''' J) Update inventories ''' 
-        market.assign_assets(pop, price)
+        market.update_inventory(pop, price)
         
         ''' K) Record results '''
         maxFitness = max(fitnessValues)

### 2021-06-07 market.py

@@ -136,8 +136,15 @@ def compute_aggregate_excess_demand(pop):
         return ((cum_sum / price) - cum_own)
     return aggregate_ed
 
-def assign_assets (pop, price):
+def update_inventory (pop, price):
     for ind in pop:
+        former_asset = ind[3]
+        former_loan = ind[4]
+        
+        # Update new asset shares
         ind[3] = ind[6] / price + ind[3]
+        
+        # Update new cash 
+        ind[2] = ind[2] - (ind[3] - former_asset) * price - ind[4] + former_loan
     return ind
 

### 2021-06-07 draft.tex
Updates from Overleaf

--- 2021-05-10 experiments.tex
@@
 \addbibresource{bibliography.bib}
+\usepackage{amsmath}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 \label{excess_demand_equation}
-   E_i(t) = \frac{W_i(t) {\lambda_i}}{p(t)} \left( \tanh{({c} \phi_i(t)}) + \frac{1}{2} \right) - S_i(t)
+   E_i(t, p(t)) = \frac{W_i(t) {\lambda_i}}{p(t)} \left( \tanh{({c} \phi_i(t)}) + \frac{1}{2} \right) - S_i(t)
 \end{equation}
@@
 when $\phi(t) = 0$), then the agent equally splits its wealth into cash and asset shares.
+
+Agent $i$'s wealth $W_i(t)$ evaluated at time $t$ is computed with:
+\begin{equation}
+    W_i(t) = C(t) + S(t)p(t) - L(t)
+\end{equation}
 
@@
 When $f < 1$, profits cause investors to withdraw funds from the agent-fund; finally, when $f=1$, no particular reinvestment dynamic occurs.
+\\
+
+Once the clearing price $p^* = \argmin \left(\sum_{a \in, \mathcal{A}} E_{a}(p)\right)^{2} $ has been determined, we update the inventories of the agents.
+First, the number of asset shares the agents own is determined as:
+
+\begin{equation}
+    S_i(t+1) = E_i(t, p^*) + S_i(t)
+\end{equation}
+
+In general, the fixed supply constraint should hold for any $t$:
+
+\begin{equation}
+    \sum_i S_i(t) = Q
+\end{equation}
+
+The new cash is:
+
+\begin{equation}
+    C(t+1) = C(t) - (S(t+1) - S(t))p(t+1) - L(t+1) + L(t)
+\end{equation}
+
+Debt
+
+% Debt is taken on when using leverage > 1 or leverage < -1.
+For long positions S(t) > 0 and we have L(t) = max(0, -(W(t-1) - S(t)p(t))).
+For short positions S(t) < 0, we have L(t) = max(0, -(W(t-1) + S(t)p(t))).
+From this, we should be able to buy W/p units without borrowing, and when short-selling we are restricted to short sell at most W/p units without borrowing.
+This is restricting the usage of the cash obtained from short-selling for something else (should we for example introduce more assets, we can not use the new cash to buy many other assets without violating leverage constraints).
+
+% Margin M(t)  is something for the balance sheet only, it tracks the current value of the debt min(0, S(t))p(t).
+When you create the short position at time t, you sell some amount of stock for cash.
+This increases the pile of cash (see equation for C(t+1)), but it increases L(t) at the same time.
+Thus it would cancel out in the wealth equation.
 

### 2021-06-07 figures/etf_time_horizons.py
demo

@@ -1,0 +1,111 @@
+import numpy as np
+import matplotlib.pyplot as plt
+class Brownian():
+    """
+    A Brownian motion class constructor
+    """
+    def __init__(self,x0=0):
+        """
+        Init class
+        """
+        assert (type(x0)==float or type(x0)==int or x0 is None), "Expect a float or None for the initial value"
+        
+        self.x0 = float(x0)
+    
+    def gen_random_walk(self,n_step=100):
+        """
+        Generate motion by random walk
+        
+        Arguments:
+            n_step: Number of steps
+            
+        Returns:
+            A NumPy array with `n_steps` points
+        """
+        # Warning about the small number of steps
+        if n_step < 30:
+            print("WARNING! The number of steps is small. It may not generate a good stochastic process sequence!")
+        
+        w = np.ones(n_step)*self.x0
+        
+        for i in range(1,n_step):
+            # Sampling from the Normal distribution with probability 1/2
+            yi = np.random.choice([1,-1])
+            # Weiner process
+            w[i] = w[i-1]+(yi/np.sqrt(n_step))
+        
+        return w
+    
+    def gen_normal(self,n_step=100):
+        """
+        Generate motion by drawing from the Normal distribution
+        
+        Arguments:
+            n_step: Number of steps
+            
+        Returns:
+            A NumPy array with `n_steps` points
+        """
+        if n_step < 30:
+            print("WARNING! The number of steps is small. It may not generate a good stochastic process sequence!")
+        
+        w = np.ones(n_step)*self.x0
+        
+        for i in range(1,n_step):
+            # Sampling from the Normal distribution
+            yi = np.random.normal()
+            # Weiner process
+            w[i] = w[i-1]+(yi/np.sqrt(n_step))
+        
+        return w
+    
+    def stock_price(
+                    self,
+                    s0=100,
+                    mu=0.2,
+                    sigma=0.68,
+                    deltaT=52,
+                    dt=0.1
+                    ):
+        """
+        Models a stock price S(t) using the Weiner process W(t) as
+        `S(t) = S(0).exp{(mu-(sigma^2/2).t)+sigma.W(t)}`
+        
+        Arguments:
+            s0: Iniital stock price, default 100
+            mu: 'Drift' of the stock (upwards or downwards), default 1
[…]

### 2021-06-07 testing_&_archiv/individual.py
leap issues

@@ -1,0 +1,236 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jun  7 15:28:51 2021
+
+@author: aymer
+"""
+
+#!/usr/bin/env python
+"""
+    Defines `Individual`
+"""
+from math import nan, isnan
+from copy import deepcopy
+from functools import total_ordering
+
+from leap_ec.decoder import IdentityDecoder
+
+
+##############################
+# Class Individual
+##############################
+@total_ordering
+class Individual:
+    """
+        Represents a single solution to a `Problem`.
+        We represent an `Individual` by a `genome` and a `fitness`.
+        `Individual` also maintains a reference to the `Problem` it will be
+        evaluated on, and an `decoder`, which defines how genomes are
+        converted into phenomes for fitness evaluation.
+    """
+
+    def __init__(self, genome, decoder=IdentityDecoder(), problem=None):
+        """
+        Initialize an `Individual` with a given genome.
+        We also require `Individual`s to maintain a reference to the `Problem`:
+        >>> from leap_ec.binary_rep.problems import MaxOnes
+        >>> from leap_ec.decoder import IdentityDecoder
+        >>> ind = Individual([0, 0, 1, 0, 1], decoder=IdentityDecoder(), problem=MaxOnes())
+        >>> ind.genome
+        [0, 0, 1, 0, 1]
+        Fitness defaults to `None`:
+        >>> ind.fitness is None
+        True
+        :param genome: is the genome representing the solution.  This can be
+            any arbitrary type that your mutation operators, probes, etc.,
+            know how to read and manipulate---a list, class, etc.
+        :param decoder: is a function or `callable` that converts a genome
+            into a phenome.
+        :param problem: is the `Problem` associated with this individual.
+        """
+        # Type checking to avoid difficult-to-debug errors
+        if isinstance(decoder, type):
+            raise ValueError(
+                f"Got the type '{decoder}' as a decoder, but expected an instance.")
+        if isinstance(problem, type):
+            raise ValueError(
+                f"Got the type '{problem}' as a problem, but expected an instance.")
+        # Core data
+        self.genome = genome
+        self.problem = problem
+        self.decoder = decoder
+        self.fitness = None
+
+    @classmethod
+    def create_population(cls, n, initialize, decoder, problem):
+        """
+        A convenience method for initializing a population of the appropriate
+        subtype.
+        :param n: The size of the population to generate
+        :param initialize: A function f(m) that initializes a genome
+        :param decoder: The decoder to attach individuals to
[…]

### 2021-06-07 testing_&_archiv/initializers.py

@@ -1,0 +1,44 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jun  7 15:31:40 2021
+
+@author: aymer
+"""
+
+#!/usr/bin/env python3
+"""
+    Initializers for real values.
+"""
+import numpy as np
+
+import individual
+
+##############################
+# Closure create_real_vector
+##############################
+def create_real_vector(bounds):
+    """
+    A closure for initializing lists of real numbers for real-valued genomes,
+    sampled from a uniform distribution.
+    Having a closure allows us to just call the returned function N times
+    in `Individual.create_population()`.
+    TODO Allow either a single tuple or a test_sequence of tuples for bounds. —Siggy
+    :param bounds: a list of (min, max) values bounding the uniform sampline
+        of each element
+    :return: A function that, when called, generates a random genome.
+    E.g., can be used for `Individual.create_population()`
+    >>> from leap_ec.decoder import IdentityDecoder
+    >>> from . problems import SpheroidProblem
+    >>> bounds = [(0, 1), (0, 1), (-1, 100)]
+    >>> population = Individual.create_population(10, create_real_vector(bounds),
+    ...                                           decoder=IdentityDecoder(),
+    ...                                           problem=SpheroidProblem())
+    """
+
+    low = [ l for l, _ in bounds ]
+    high = [ h for _, h in bounds ]
+
+    def create():
+        return np.random.uniform(low, high, size=len(bounds))
+
+    return create

### 2021-06-07 testing_&_archiv/market_clearing_leap.py

@@ -1,24 +1,15 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Jun  7 08:38:08 2021
-
-@author: aymer
-"""
-
 """
     Provides a very high-level convenience function for a very general EA,
     ea_solve().
 """
 from matplotlib import pyplot as plt
-
+import individual
+import representation
 from leap_ec import ops, probe
 from leap_ec.algorithm import generational_ea
-from leap_ec.real_rep.ops import mutate_gaussian
-from leap_ec.representation import Representation
-from leap_ec.individual import Individual
 from leap_ec.problem import FunctionProblem
-from leap_ec.decoder import IdentityDecoder
-from leap_ec.real_rep.initializers import create_real_vector
+import initializers
+from leap_ec.real_rep.ops import mutate_gaussian
 
 
 ##############################
@@ -85,13 +76,13 @@ def ea_solve(function, bounds, generations=100, pop_size=2,
         plot_probe = probe.FitnessPlotProbe(ylim=viz_ylim, ax=plt.gca())
         pipeline.append(plot_probe)
 
-    ea = generational_ea(generations=generations, pop_size=pop_size,
+    ea = generational_ea(max_generations=generations,
+                         pop_size=pop_size,
                          problem=FunctionProblem(function, maximize),
 
-                         representation=Representation(
-                             individual_cls=Individual,
-                             decoder=IdentityDecoder(),
-                             initialize=create_real_vector(bounds=bounds)
+                         representation=representation.Representation(
+                             individual_cls=individual,
+                             initialize=initializers.create_real_vector(bounds=bounds)
                          ),
 
                          pipeline=pipeline)
@@ -107,4 +98,13 @@ def ea_solve(function, bounds, generations=100, pop_size=2,
 
 def f(x):
     return sum(x)**2
-ea_solve(f, bounds=[(4, 5.12) for i in range(1)], maximize=False, hard_bounds = True)
+ea_solve(f, bounds=[(4, 5.12) for i in range(1)], maximize=False, hard_bounds = True, mutation_std = 0)
+
+
+from leap_ec.simple import ea_solve
+
+def f(x):
+    """A real-valued function to optimized."""
+    return sum(x)**2
+
+ea_solve(f, bounds=[(4, 5.12) for _ in range(1)], maximize=False, mutation_std = 0, pop_size = 100)

### 2021-06-07 testing_&_archiv/representation.py

@@ -1,0 +1,47 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jun  7 15:30:09 2021
+
+@author: aymer
+"""
+
+#!/usr/bin/env python3
+"""
+A `Representation` is a simple data structure that wraps the
+components needed to define, initialize, and decode individuals.
+This just serves as some syntactic sugar when we are specifying
+algorithms---so that representation-related components are grouped
+together and clearly labeled `Representation`.
+"""
+from leap_ec.individual import Individual
+from leap_ec.decoder import IdentityDecoder
+
+##############################
+# Class Representation
+##############################
+class Representation():
+    """ Syntactic sugar for some of the monolithic functions that
+        conveniently combines a decoder, initializer, and an Individual
+        class since those always work in tandem, but can still be loosely
+        coupled.
+     """
+
+    def __init__(self, initialize, decoder=IdentityDecoder(), individual_cls=Individual):
+        self.decoder = decoder
+        self.initialize = initialize
+        self.individual_cls = individual_cls
+
+    def create_individual(self, problem):
+        """Make a single individual."""
+        return self.create_population(1, problem)[0]
+
+    def create_population(self, pop_size, problem):
+        """ make a new population
+        :param pop_size: how many individuals should be in the population
+        :param problem: to be solved
+        :return: a population of `individual_cls` individuals
+        """
+        return self.individual_cls.create_population(pop_size,
+                                                     initialize=self.initialize,
+                                                     decoder=self.decoder,
+                                                     problem=problem)

### 2021-06-08 draft.tex
Updates from Overleaf

@@
 \usepackage{hyperref}
+\usepackage{xcolor}
 \newcommand{\code}{\texttt}
@@
 The $\theta$ parameter describes the time horizon on which the trend following strategy is based (Equation \ref{trend_following_equation}).
-Initial strategies are drawn uniformly in $(1,H]$, where $H$ is the maximum possible time horizon.
+Initial strategies are drawn uniformly in $[2,H]$, where $H$ is the maximum possible time horizon.
 \par
@@
 \begin{equation}
-    W_i(t) = C(t) + S(t)p(t) - L(t)
+    W_i(t) = C_i(t) + S_i(t)p(t) - L_i(t)
 \end{equation}
@@
 
-This process allows to pin down the clearing price $p(t)$.
+This process allows to pin down the clearing price $p^* = p(t+1)$.
 We can thus compute the realised demands of the agents, update their asset shares ownership, wealth and profit.
@@
 
-Debt
+\textbf{Margin}
+
+When the agent enters into a short position (i.e.
+$S(t+1) = E(t,p(t+1)) - S(t) < 0$ , it sells asset shares that it does not own.
+Its cash grows by $p(t+1)S(t+1)$, but the agent enters into debt, as it must buy later $S(t+1)$ assets to return them to the lender.
+
+\textcolor{red}{WFF}
+
+% In order to guarantee that the short-selling fund can
+% return the borrowed securities to the lender at a later time, the fund
+% sets aside a margin amount M equal to the current market value
+% of the borrowings, in the form of cash.
+
+% - Why the current values? Would it make sense for the agent to remember that it must return S securities, regardless of their price?
+% - If the margin amount is set aside in the form of cash, Cash increases by $p(t+1)E(t,p(t+1))$ but decreases by $p(t+1)E(t,p(t+1))$ to fuel the Margin, hence no benefit was done?
+(unless we automatically return the borrowed assets in the next period at a lower price?)
+% - How is returning borrowed securities handled?
+Does this happen automatically at the next period after short selling?
+Related question: as Margin does not appear in the wealth equation, how is it impacting the agent?
 
@@
 This models funds, managers, who look at the profits and performance of other funds, and tend to imperfectly imitate stronger financial strategies.
-Each fund can also independently choose to marginally change its strategy, in an exploratory behavior.\\
+Each fund can also independently choose to marginally change its strategy, in an exploratory behavior.
+First, insolvent funds cannot influence the evolution of others, and are withdrawn from the market with the \textit{hypermutation} operator.
+Then, during \textit{selection}, funds choose some other funds to compare their profits and strategy with.
+\textit{Fitness} defines the objective used for this comparison.
+Combining the strategy of the best observed fund with theirs happens during \textit{crossover}.
+Finally, \textit{mutation} allows funds to slightly adjust their strategy as small exploratory jumps.\\
 % some refs needed here to support that this behavior exists and is well documented
 
-First, we implement an \textit{hypermutation} operator in which insolvent agents are removed from the populations.
+\paragraph{Hypermutation} First, we implement an \textit{hypermutation} operator in which insolvent agents are removed from the populations.
 Agents with wealth $W \leq 0$ disappear from the market, and are replaced by new individuals with random strategies and usual initial endowments.
-% they may have to be initialised only with cash because otherwise we are creating new asset shares ex nihilo.
 
@@
 
-In this model, this imitation and exploration behavior are modeled using a genetic algorithm.
+\paragraph{Selection} In this model, this imitation and exploration behavior are modeled using a genetic algorithm.
 We first operate a modified tournament \textit{selection} with probability $\varsigma$, the step in which each fund selects a few funds it will compare itself to.
@@
 
-Once imitation by tournament selection is operated, we allow strategies to recombine together and explore alternative regions of the strategy space with probability $\rho$.
+\paragraph{Crossover} Once imitation by tournament selection is operated, we allow strategies to recombine together and explore alternative regions of the strategy space with probability $\rho$.
 This is the \textit{crossover} operator.
@@
 
-Finally, marginal exploration behavior is allowed by the \textit{mutation} operator. With probability $\mu$,
+\paragraph{Mutation} Finally, marginal exploration behavior is allowed by the \textit{mutation} operator.
+With probability $\mu$, the strategy $\theta$ of the agent changes by $+1$ or $-1$, with equal probabilities.
+The mutation operator is constrained such that the mutated strategy stays within the feasible strategies $[2, H]$.
 

### 2021-06-08 draft.tex
Updates from Overleaf

@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\usepackage{algorithm}% http://ctan.org/pkg/algorithms 
+\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx 
+\newcommand{\var}[1]{{\ttfamily#1}}% variable
 
@@
 
-\section{Preamble: baseline configuration}
+% Necessary condition: running GA with complete market structure
+% To implement:
+% \begin{itemize}
+%     \item Observe market dynamics over time (easy): price, returns, assets, volatility, dividends, profits
+%     \item Observe ecology dynamics over time (hard): proportions? heatmap? Add entropy?
+%     Perhaps H (for max time horizon) plots with clear color codes (like jet) over time, and hopefully we see something clear happen.
+%     \item Develop a GUI or something nice/updating itself over time to observe that?
+% \end{itemize}
 
-Necessary condition: running GA with complete market structure
-Objective: be as close as possible to Maarten's paper configuration, notably for market environment parameters.
-Add Agent leverage, aggressiveness of response to be closer to Maarten
-\section{Preamble 2: observables}
+% \section{Trend follower market ecology}
 
-Necessary condition: running GA with complete market structure
-To implement:
-\begin{itemize}
-    \item Observe market dynamics over time (easy): price, returns, assets, volatility, dividends, profits
-    \item Observe ecology dynamics over time (hard): proportions? heatmap? Add entropy?
-    Perhaps H (for max time horizon) plots with clear color codes (like jet) over time, and hopefully we see something clear happen.
-    \item Develop a GUI or something nice/updating itself over time to observe that?
-\end{itemize}
+% Necessary condition: a running GA with a single asset, market clearing, and adequate starting conditions, and the baseline configuration.
 
-\section{Trend follower market ecology}
-
-Necessary condition: a running GA with a single asset, market clearing, and adequate starting conditions, and the baseline configuration.
-
-\subsection{Influence of learning vs non-learning}
 The non-lerning case (replacement dynamics only) can be illustrated with:

### 2021-06-08 data.py
step 1 towards df output

@@ -1,0 +1,8 @@
+''' Here we consider the variables we recorded during the run, and we assemble 
+them into a coherent dataframe for later statistical and graphical analysis'''
+
+import pandas as pd
+
+def generate_df():
+    df = 0 #temp
+    return df

### 2021-06-08 main.py

@@ -6,6 +6,7 @@ import genetic_algorithm_functions as ga
 import market as market
 # import market_clearing as mc
 import parameters
+import data
 
 RANDOM_SEED = parameters.RANDOM_SEED
 POPULATION_SIZE = parameters.POPULATION_SIZE
@@ -173,5 +174,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         #     for ind in pop:
         #         ind[1] -= 1
 
+        # XYZ = data.generate_df()
+    # return 
     
     return initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed

### 2021-06-08 run.py

@@ -1,0 +1,74 @@
+import matplotlib.pyplot as plt
+import main
+import parameters
+MAX_GENERATIONS = parameters.MAX_GENERATIONS
+
+# Optional because they are arguments to main
+CROSSOVER_RATE = parameters.CROSSOVER_RATE
+MUTATION_RATE = parameters.MUTATION_RATE 
+
+
+initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed = main.main(0, CROSSOVER_RATE, MUTATION_RATE)
+
+# Plot population histograms at the start and at the end
+print("--------------------------")
+
+# print("Initial population was " + str(initial_pop))
+# sns.histplot(data=np.array(initial_pop), legend = False, stat = "density", shrink = 0.85, discrete=True, bins = 11)
+# plt.show()
+# print("Current population is " + str(pop))
+# sns.histplot(data=np.array(pop), legend = False, stat = "density", shrink = 0.85, discrete=True, bins = 11)
+# plt.show()
+
+# Plot the fitness evolution over time
+plt.plot(maxFitnessValues, color='red', label='Maximum fitness')
+plt.plot(meanFitnessValues, color='green', label = 'Average fitness')
+plt.xlabel('Generations')
+plt.ylabel('Max / Average Fitness')
+plt.title('Max and Average Fitness over Generations')
+plt.xlim(0,MAX_GENERATIONS+1)
+plt.legend()
+plt.show()
+
+plt.plot(replacements, color='gray', label = 'Hypermutations')
+plt.xlabel('Generations')
+plt.ylabel('Hypermutations')
+plt.title('Hypermutations over time')
+plt.xlim(0,MAX_GENERATIONS+1)
+plt.legend()
+plt.show()
+
+plt.plot(agent0_ema, color='red', label='EMA')
+plt.plot(agent0_profit, color='green', label = 'Profits')
+plt.xlabel('Generations')
+plt.ylabel('Profit')
+plt.title('Profit and EMA profit of Albert')
+plt.xlim(0,MAX_GENERATIONS+1)
+plt.legend()
+plt.show()
+
+
+plt.plot(agent0_ema, color='red', label='EMA')
+plt.plot(agent0_profit, color='green', label = 'Profits')
+plt.xlabel('Generations')
+plt.ylabel('Profit')
[…]

### 2021-06-08 data.py
adding some df, correcting some mismatch in list length

@@ -1,8 +1,32 @@
 ''' Here we consider the variables we recorded during the run, and we assemble 
 them into a coherent dataframe for later statistical and graphical analysis'''
 
+''' We want data on:
+    
+--- Economic variables ---
+- Price series
+
+--- Ecology variables ---
+
+'''
+
+#  One large dataframe with all, or separate? TBD
+# We can also adjust the variables order in the columns
+
 import pandas as pd
 
-def generate_df():
-    df = 0 #temp
+def generate_df(generation_history, price_history, dividend_history, replacements):
+    df = pd.DataFrame()
+    
+    # General variables
+    
+    df["Gen"] = generation_history
+    
+    # Economic variables 
+    df["Price"] = price_history
+    df["Div"] = dividend_history
+    
+    # Ecology variables
+    df ["Rep"] = replacements
+    
     return df

### 2021-06-08 main.py

@@ -36,6 +36,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     dividend_history = []
     random_dividend_history = []
     price_history = []
+    generation_history = []
 
     
     # Temp
@@ -43,7 +44,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     agent0_ema = []
     
     dividend = INITIAL_DIVIDEND
-    dividend_history.append(INITIAL_DIVIDEND)
+    # dividend_history.append(INITIAL_DIVIDEND)
     
     print("Initial population")
     print(('{}\n'*len(pop)).format(*pop))
@@ -60,7 +61,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     meanFitness = sum(fitnessValues) / len(pop)
     maxFitnessValues.append(maxFitness)
     meanFitnessValues.append(meanFitness)
-    replacements.append(0)
+    # replacements.append(0)
     # Temp
     agent0_profit.append(pop[0][7])
     agent0_ema.append(pop[0][8])
@@ -167,6 +168,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         #  Could this print results be automated? We have it twice
         
         print("- Generation {}: Max Fitness = {}, Avg Fitness = {}".format(generationCounter, maxFitness, meanFitness))
+        generation_history.append(generationCounter)
         generationCounter += 1
                 
         # Temporary function to apply some fixed cost
@@ -174,7 +176,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         #     for ind in pop:
         #         ind[1] -= 1
 
-        # XYZ = data.generate_df()
+        df = data.generate_df(generation_history, price_history, dividend_history, replacements)
     # return 
     
-    return initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed
+    return initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df

### 2021-06-08 market.py

@@ -56,7 +56,7 @@ def draw_dividend(DIVIDEND_GROWTH_RATE):
     # wiener.append(random_dividend)
     
     dividend = abs(dividend + DIVIDEND_GROWTH_RATE * dividend + DIVIDEND_GROWTH_VOLATILITY * dividend * random_dividend)
-    
+    dividend = truncate(dividend, 3)
     return dividend, random_dividend
 
         

### 2021-06-08 run.py

@@ -8,8 +8,8 @@ CROSSOVER_RATE = parameters.CROSSOVER_RATE
 MUTATION_RATE = parameters.MUTATION_RATE 
 
 
-initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed = main.main(0, CROSSOVER_RATE, MUTATION_RATE)
-
+initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df = main.main(0, CROSSOVER_RATE, MUTATION_RATE)
+print(df)
 # Plot population histograms at the start and at the end
 print("--------------------------")
 

### 2021-06-08 data.py
df update, looking good

@@ -1,12 +1,19 @@
 ''' Here we consider the variables we recorded during the run, and we assemble 
 them into a coherent dataframe for later statistical and graphical analysis'''
 
-''' We want data on:
+''' We want data on: [TO COMPLETE AS WE GO]
+
+--- General ---
+- Generation index
     
 --- Economic variables ---
-- Price series
+- Price 
+- Dividend 
+- Random dividend component 
+
 
 --- Ecology variables ---
+- Replacements 
 
 '''
 
@@ -15,7 +22,7 @@ them into a coherent dataframe for later statistical and graphical analysis'''
 
 import pandas as pd
 
-def generate_df(generation_history, price_history, dividend_history, replacements):
+def generate_df(generation_history, price_history, dividend_history, random_dividend_history, replacements):
     df = pd.DataFrame()
     
     # General variables
@@ -25,6 +32,7 @@ def generate_df(generation_history, price_history, dividend_history, replacement
     # Economic variables 
     df["Price"] = price_history
     df["Div"] = dividend_history
+    df["RDiv"] = random_dividend_history
     
     # Ecology variables
     df ["Rep"] = replacements

### 2021-06-08 main.py

@@ -176,7 +176,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         #     for ind in pop:
         #         ind[1] -= 1
 
-        df = data.generate_df(generation_history, price_history, dividend_history, replacements)
+        df = data.generate_df(generation_history, price_history, dividend_history, random_dividend_history, replacements)
     # return 
     
     return initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df

### 2021-06-08 market.py

@@ -50,7 +50,7 @@ def draw_dividend(DIVIDEND_GROWTH_RATE):
     global dividend
     global random_dividend
     
-    random_dividend = random.normalvariate(0, 1)
+    random_dividend = truncate(random.normalvariate(0, 1),3)
     if len(random_dividend_history) > DIVIDEND_ATC_TAU:
         random_dividend = (1 - DIVIDEND_AUTOCORRELATION ** 2) * random_dividend + DIVIDEND_AUTOCORRELATION * random_dividend_history[len(random_dividend_history) - 1 - DIVIDEND_ATC_TAU]
     # wiener.append(random_dividend)

### 2021-06-09 market_clearing_esl.py
failed attempts on clearing

@@ -1,3 +1,52 @@
+# import esl
+# from esl.law import property
+# from esl.simulation import identity
+# from esl.economics import price
+# from esl.economics import currencies
+# from esl.economics.markets import quote
+# from esl.economics.markets.walras import excess_demand_model, differentiable_order_message
+
+# def solve(excess_demand_functions: list):
+#     i  = identity([0, 1])
+#     p  = property(i)
+#     initial_price = price(int(1.23 * currencies.USD.denominator), currencies.USD)
+#     initial_quote = quote(initial_price)
+#     market_agent = esl.simulation.identity([0])
+#     excess_demand_functions = []
+#     model = excess_demand_model({p: initial_quote})
+
+#     # for i, ed in enumerate(my_excess_demand_functions):
+
+#     #     order = ed(esl.simulation.identity([i+1]), market_agent, 0, 0)
+
+#     #     excess_demand_functions.append(order) 
+
+#     model.excess_demand_functions = excess_demand_functions
+#     multipliers = model.compute_clearing_quotes() 
+
+#     prices = []
+
+#     for k, v in multipliers.items():
+
+#         prices.append(price(round(float(initial_price) * v * currencies.USD.denominator), currencies.USD))
+
+
+#     del model
+#     return prices
+
+# def aggregate_ed1(x):
+#     return 3 * x - 10
+
+# def aggregate_ed2(x):
+#     return 5 * x - 7
+
+# aggregate_ed = [aggregate_ed1, aggregate_ed2]
+
+# prices = solve(aggregate_ed)
+# print(prices)
+
+''' As understood on 09/06: ed1, ed2 would have to be orders, not functions '''
+
 import esl
 from esl.law import property
 from esl.simulation import identity
@@ -6,52 +55,20 @@ from esl.economics import currencies
 from esl.economics.markets import quote
 from esl.economics.markets.walras import excess_demand_model, differentiable_order_message
 
-def solve(excess_demand_functions: list):
-    i  = identity([0, 1])
-    p  = property(i)
-    initial_price = price(int(1.23 * currencies.USD.denominator), currencies.USD)
-    initial_quote = quote(initial_price)
-    market_agent = esl.simulation.identity([0])
-    excess_demand_functions = []
-    model = excess_demand_model({p: initial_quote})
+class my_order(differentiable_order_message):
+    def __init__(self, sender, recipient, sent, received):
+        super().__init__(sender, recipient, sent, received)
[…]

### 2021-06-09 genetic_algorithm_functions.py
added margin inventory

@@ -47,11 +47,12 @@ toolbox.register("generate_trading_signal", random.randint, 0, 0)
 toolbox.register("generate_excess_demand", random.randint, 0, 0)
 toolbox.register("generate_profit", random.randint, 0, 0)
 toolbox.register("generate_ema", random.randint, 0, 0)
+toolbox.register("generate_margin", random.randint, 0, 0)
 
 toolbox.register("generate_individual", tools.initCycle, creator.individual, 
                  (toolbox.generate_strategy, toolbox.generate_wealth, toolbox.generate_cash, 
                   toolbox.generate_asset, toolbox.generate_loan, toolbox.generate_trading_signal, 
-                  toolbox.generate_excess_demand,toolbox.generate_profit,toolbox.generate_ema), n=1)
+                  toolbox.generate_excess_demand,toolbox.generate_profit,toolbox.generate_ema, toolbox.generate_margin), n=1)
 toolbox.register("population_creation", tools.initRepeat, list, toolbox.generate_individual)
 
 '''
@@ -121,7 +122,7 @@ def selTournament(individuals, k, tournsize, fit_attr="fitness"):
         aspirants = selRandom(individuals, tournsize-1) 
         aspirants.append(individuals[i])
         chosen_i = max(aspirants, key=attrgetter(fit_attr))
-        chosen_i[1:9] = individuals[i][1:9]
+        chosen_i[1:10] = individuals[i][1:10]
         chosen.append(chosen_i)
     return chosen
 

### 2021-06-09 main.py

@@ -49,8 +49,8 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     print("Initial population")
     print(('{}\n'*len(pop)).format(*pop))
     print("Agent representation")
-    print("[Theta Wealth Cash Asset Loan TradingSignal RawExcessDemand     Profit     EMA profit]")
-    print("[ 0       1     2    3     4         5             6           7            8 ]")
+    print("[Theta Wealth Cash Asset Loan TradingSignal RawExcessDemand     Profit     EMA profit   Margin]")
+    print("[ 0       1     2    3     4         5             6           7            8              9]")
     
     fitnessValues = list(map(ga.toolbox.evaluate, pop))
     for individual, fitnessValue in zip(pop, fitnessValues):

### 2021-06-09 genetic_algorithm_functions.py
correct hypermutation
To not classify as insolvent agents in a short position with a sufficient margin

@@ -135,7 +135,7 @@ def hypermutate(pop):
     pop_temp = list(map(toolbox.clone, pop))
     
     for i in range(0, len(pop_temp)):
-        if pop_temp[i][1] <= 0:
+        if pop_temp[i][1] + pop_temp[i][9] <= 0:
             pop_temp[i] = toolbox.generate_individual()
             del pop_temp[i].fitness.values
             global round_replacements

### 2021-06-09 market.py

@@ -140,11 +140,21 @@ def update_inventory (pop, price):
     for ind in pop:
         former_asset = ind[3]
         former_loan = ind[4]
+        realised_ed = ind[6] / price + ind[3]
         
         # Update new asset shares
-        ind[3] = ind[6] / price + ind[3]
+        ind[3] = realised_ed
+        
+        # If short selling, set some margin aside
+        if ind[3] < 0:
+            ind[9] += ind[3] * price
         
         # Update new cash 
-        ind[2] = ind[2] - (ind[3] - former_asset) * price - ind[4] + former_loan
+        ind[2] = ind[2] - (ind[3] - former_asset) * price - ind[4] + former_loan - ind[9]
+        
+        #	 Clear the margin if we are out of the short position
+        if ind[3] >= 0:
+            ind[2] += ind[9]
+            ind[9] = 0
     return ind
 

### 2021-06-09 market.py
improvements

@@ -104,13 +104,8 @@ def update_trading_signal(pop, price_history):
     return ind
 
 def update_excess_demand(pop):
-    # global list_excess_demand_func
-    
     for ind in pop:
         ind[6] = ind[1] * LAMBDA_TF * (np.tanh(STRATEGY_AGGRESSIVENESS_TF * ind[5]) + 0.5)
-    # ''' This will need to add the division by the price '''
-    # ''' This will probably have to be translated into a function '''
-
     return ind
 
 def order_excess_demand(pop):
@@ -122,11 +117,6 @@ def order_excess_demand(pop):
     return list_excess_demand_func
 
 def compute_aggregate_excess_demand(pop):
-    # def aggregate_ed(price):
-    #     return 0
-    # for i in range(len(list_excess_demand_func)):
-    #     def aggregate_ed(price):
-    #         return aggregate_ed(price) + list_excess_demand_func[i](price)
     cum_sum = 0
     cum_own = 0
     for ind in pop:

### 2021-06-14 genetic_algorithm_functions.py
important bug fix

@@ -28,8 +28,8 @@ INITIAL_CASH = parameters.INITIAL_CASH
 # =============================================================================
 
 # Agent representaiton:
-#     [Theta Wealth Cash Asset Loan TradingSignal ExcessDemand     Profit     EMA profit]
-#     [ 0       1     2    3     4         5             6           7            8 ]
+#     [Theta Wealth Cash Asset Loan TradingSignal ExcessDemand     Profit     EMA profit   Margin]
+#     [ 0       1     2    3     4         5             6           7            8           9]
 
 toolbox = base.Toolbox()
 

### 2021-06-14 main.py

@@ -133,19 +133,27 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         market.update_excess_demand(pop)
         list_excess_demand_func = market.order_excess_demand(pop)
         ''' list_excess_demand_func is now the list of ED functions '''
-        aggregate_ed = market.compute_aggregate_excess_demand(pop)
+        # aggregate_ed = market.compute_aggregate_excess_demand(pop)
+        aggregate_ed = market.compute_aggregate_excess_demand(pop, list_excess_demand_func)
         
         print("After trading signal, GA, ED update, right before clearing")
         print(('{}\n'*len(pop)).format(*pop))
 
         ''' I) Clear the market with the aggregate ED, obtain the new price ''' 
-        # price = mc_leap.leap_solver(aggregate_ed, price)
-        # print(price)
-        # print(aggregate_ed(price))
-        price = mc_leap.solver_linear_shortcut(pop, price)
+        
+        price = mc_leap.leap_solver(pop, price)
         print(price)
         print(aggregate_ed(price))
-         
+        # price = mc_leap.solver_linear_shortcut(pop, price)
+        # print(price)
+        # print(aggregate_ed(price))
+        
+        # testing block
+        sum_ag = 0
+        for i in range(POPULATION_SIZE):
+            sum_ag += list_excess_demand_func[i](price)
+            print("Agent " + str(i) + " ED(pt+1) is " + str(list_excess_demand_func[i](price)))
+        print("Sum AG " + str(sum_ag))
         price_history.append(price)
         
         ''' J) Update inventories ''' 
@@ -176,4 +184,4 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
                               dividend_history, random_dividend_history, replacements)
     # return 
     
-    return initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df
+    return price, initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df

### 2021-06-14 market.py

@@ -111,26 +111,39 @@ def update_excess_demand(pop):
 def order_excess_demand(pop):
     list_excess_demand_func = []
     for ind in pop:
-        def ed(price):
-            return ind[6] / price - ind[3]
+        
+        # Testing block
+        
+        
+        def ed(x):
+            
+            return truncate((ind[6] / x) - ind[3],4)
         list_excess_demand_func.append(ed)
+        print(ed(1))
+        del ed
     return list_excess_demand_func
 
-def compute_aggregate_excess_demand(pop):
-    cum_sum = 0
-    cum_own = 0
-    for ind in pop:
-        cum_sum += ind[6]
-        cum_own += ind[3]
-    def aggregate_ed(price):
-        return ((cum_sum / price) - cum_own)
+def compute_aggregate_excess_demand(pop, list_excess_demand_func):
+    def aggregate_ed(x):
+        result = 0
+        for i in range(len(pop)):
+            result += list_excess_demand_func[i](x)
+        return result
     return aggregate_ed
+    # cum_sum = 0
+    # cum_own = 0
+    # for ind in pop:
+    #     cum_sum += ind[6]
+    #     cum_own += ind[3]
+    # def aggregate_ed(price):
+    #     return ((cum_sum / price) - cum_own)
+    # return aggregate_ed
 
 def update_inventory (pop, price):
     for ind in pop:
         former_asset = ind[3]
         former_loan = ind[4]
-        realised_ed = truncate(ind[6] / price + ind[3],3)
+        realised_ed = truncate(ind[6] / price + ind[3],4)
         
         
         ''' If we want to buy assets: non-negative cash buying procedure '''

### 2021-06-14 market_clearing_leap.py

@@ -44,15 +44,26 @@ def ea_solve_noverbose(function, bounds, generations=100, pop_size=2,
     return best_genome
 
 
-def leap_solver(aggregate_ed, price):
-    ''' First define squared aggregate demand '''
+def leap_solver(pop, current_price):
+    
+    ''' Define our workable aggregate demands '''
+    cum_sum = 0
+    cum_own = 0
+    for ind in pop:
+        cum_sum += ind[6]
+        cum_own += ind[3]
+    def agg_ed_solver(x):
+        return cum_sum / sum(x) - cum_own
+    
+    
+    ''' define squared aggregate demand '''
     def squared_agg_ed(x):
-        result = aggregate_ed(x) ** 2
+        result = agg_ed_solver(x) ** 2
         return result
     
     ''' Then define the circuit breaker bounds '''
-    limit_down = price * 0.5
-    limit_up = price * 2.0
+    limit_down = current_price * 0.5
+    limit_up = current_price * 2.0
     
     ''' Run the solver '''
     best_genome = ea_solve_noverbose(squared_agg_ed,
@@ -63,6 +74,27 @@ def leap_solver(aggregate_ed, price):
     ''' TODO: replace this clearing by the better ESL solver '''
     return(market.truncate(best_genome[0],3))
 
+# def aggregate_ed(x):
+#     return 1000 / sum(x) + 5013
+
+# leap_solver(aggregate_ed, 3)
+# limit_down = 0
+# limit_up = 10
+# best_genome = ea_solve_noverbose(aggregate_ed, bounds=[(limit_down, limit_up) for _ in range(1)], 
+#                                  generations = 50, pop_size = 500, mutation_std=0, maximize = False)
+
+
+
+# def f(x):
+#     """A real-valued function to optimized."""
+#     return sum(x)**2
+# def g(x):
+#     """A real-valued function to optimized."""
+#     return x / 2 + 10
+# from leap_ec.simple import ea_solve
+# ea_solve(f, bounds=[(-5.12, 5.12) for _ in range(5)], maximize=True)
+# ea_solve(g, bounds=[(-5.12, 5.12) for _ in range(5)], maximize=True)
+
 def solver_linear_shortcut(pop, price):
     cum_sum = 0
     cum_own = 0
@@ -72,14 +104,24 @@ def solver_linear_shortcut(pop, price):
         
     candidate = (cum_sum / cum_own)
     
-    limit_down = price * 0.5
-    limit_up = price * 2.0
[…]

### 2021-06-14 run.py

@@ -8,7 +8,7 @@ CROSSOVER_RATE = parameters.CROSSOVER_RATE
 MUTATION_RATE = parameters.MUTATION_RATE 
 PROBA_SELECTION = parameters.PROBA_SELECTION
 
-initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df = main.main(PROBA_SELECTION, CROSSOVER_RATE, MUTATION_RATE)
+price, initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df = main.main(PROBA_SELECTION, CROSSOVER_RATE, MUTATION_RATE)
 print(df)
 # Plot population histograms at the start and at the end
 print("--------------------------")

### 2021-06-14 market.py
cleaning

@@ -111,9 +111,7 @@ def update_excess_demand(pop):
 def order_excess_demand(pop):
     list_excess_demand_func = []
     for ind in pop:
-        
-        # Testing block
-        
+
         
         def ed(x):
             
@@ -130,14 +128,7 @@ def compute_aggregate_excess_demand(pop, list_excess_demand_func):
             result += list_excess_demand_func[i](x)
         return result
     return aggregate_ed
-    # cum_sum = 0
-    # cum_own = 0
-    # for ind in pop:
-    #     cum_sum += ind[6]
-    #     cum_own += ind[3]
-    # def aggregate_ed(price):
-    #     return ((cum_sum / price) - cum_own)
-    # return aggregate_ed
+
 
 def update_inventory (pop, price):
     for ind in pop:

### 2021-06-14 market_clearing_leap.py

@@ -74,26 +74,7 @@ def leap_solver(pop, current_price):
     ''' TODO: replace this clearing by the better ESL solver '''
     return(market.truncate(best_genome[0],3))
 
-# def aggregate_ed(x):
-#     return 1000 / sum(x) + 5013
 
-# leap_solver(aggregate_ed, 3)
-# limit_down = 0
-# limit_up = 10
-# best_genome = ea_solve_noverbose(aggregate_ed, bounds=[(limit_down, limit_up) for _ in range(1)], 
-#                                  generations = 50, pop_size = 500, mutation_std=0, maximize = False)
-
-
-
-# def f(x):
-#     """A real-valued function to optimized."""
-#     return sum(x)**2
-# def g(x):
-#     """A real-valued function to optimized."""
-#     return x / 2 + 10
-# from leap_ec.simple import ea_solve
-# ea_solve(f, bounds=[(-5.12, 5.12) for _ in range(5)], maximize=True)
-# ea_solve(g, bounds=[(-5.12, 5.12) for _ in range(5)], maximize=True)
 
 def solver_linear_shortcut(pop, price):
     cum_sum = 0
@@ -104,24 +85,16 @@ def solver_linear_shortcut(pop, price):
         
     candidate = (cum_sum / cum_own)
     
-    # limit_down = price * 0.5
-    # limit_up = price * 2.0
+    limit_down = price * 0.5
+    limit_up = price * 2.0
     
-    # if candidate < limit_down:
-    #     candidate = limit_down
-    # if candidate > limit_up:
-    #     candidate = limit_up
+    if candidate < limit_down:
+        candidate = limit_down
+    if candidate > limit_up:
+        candidate = limit_up
         
     new_price = candidate
         
     return new_price 
 
 
-
-# from leap_ec.simple import ea_solve
-
-# def f(x):
-#     """A real-valued function to optimized."""
-#     return sum(x)**2
-
-# ea_solve(f, bounds=[(-5.12, 5.12) for _ in range(5)], maximize=True, hard_bounds = True)

### 2021-06-15 main.py
df output 1.0

@@ -184,4 +184,5 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
                               dividend_history, random_dividend_history, replacements)
     # return 
     
-    return price, initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df
+    # return price, initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df
+    return df

### 2021-06-15 run.py

@@ -8,7 +8,9 @@ CROSSOVER_RATE = parameters.CROSSOVER_RATE
 MUTATION_RATE = parameters.MUTATION_RATE 
 PROBA_SELECTION = parameters.PROBA_SELECTION
 
-price, initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df = main.main(PROBA_SELECTION, CROSSOVER_RATE, MUTATION_RATE)
+# price, initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df = main.main(PROBA_SELECTION, CROSSOVER_RATE, MUTATION_RATE)
+df = main.main(PROBA_SELECTION, CROSSOVER_RATE, MUTATION_RATE)
+
 print(df)
 # Plot population histograms at the start and at the end
 print("--------------------------")
@@ -21,54 +23,54 @@ print("--------------------------")
 # plt.show()
 
 # Plot the fitness evolution over time
-plt.plot(maxFitnessValues, color='red', label='Maximum fitness')
-plt.plot(meanFitnessValues, color='green', label = 'Average fitness')
-plt.xlabel('Generations')
-plt.ylabel('Max / Average Fitness')
-plt.title('Max and Average Fitness over Generations')
-plt.xlim(0,MAX_GENERATIONS+1)
-plt.legend()
-plt.show()
+# plt.plot(maxFitnessValues, color='red', label='Maximum fitness')
+# plt.plot(meanFitnessValues, color='green', label = 'Average fitness')
+# plt.xlabel('Generations')
+# plt.ylabel('Max / Average Fitness')
+# plt.title('Max and Average Fitness over Generations')
+# plt.xlim(0,MAX_GENERATIONS+1)
+# plt.legend()
+# plt.show()
 
-plt.plot(replacements, color='gray', label = 'Hypermutations')
-plt.xlabel('Generations')
-plt.ylabel('Hypermutations')
-plt.title('Hypermutations over time')
-plt.xlim(0,MAX_GENERATIONS+1)
-plt.legend()
-plt.show()
+# plt.plot(replacements, color='gray', label = 'Hypermutations')
[…]

### 2021-06-15 main.py
important error solved in update inventory

@@ -143,12 +143,14 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         
         price = mc_leap.leap_solver(pop, price)
         print(price)
+        print("agg_ed(price)")
         print(aggregate_ed(price))
         # price = mc_leap.solver_linear_shortcut(pop, price)
         # print(price)
         # print(aggregate_ed(price))
         
         # testing block
+        
         sum_ag = 0
         for i in range(POPULATION_SIZE):
             sum_ag += list_excess_demand_func[i](price)
@@ -165,6 +167,10 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         maxFitnessValues.append(maxFitness)
         meanFitnessValues.append(meanFitness)
         replacements.append(round_replacements)
+        
+        print("agg_price before mismatch")
+        print(aggregate_ed(price))
+        print(abs(market.truncate(aggregate_ed(price),3)))
         mismatch_history.append(abs(market.truncate(aggregate_ed(price),3)))
                 # Temp
         agent0_profit.append(pop[0][7])

### 2021-06-15 market.py

@@ -134,7 +134,9 @@ def update_inventory (pop, price):
     for ind in pop:
         former_asset = ind[3]
         former_loan = ind[4]
-        realised_ed = truncate(ind[6] / price + ind[3],4)
+        realised_ed = truncate(ind[6] / price - ind[3],4)
+        realised_ed = ind[6] / price - ind[3]
+        print("agent wants up to " + str(realised_ed))
         
         
         ''' If we want to buy assets: non-negative cash buying procedure '''

### 2021-06-15 parameters.py

@@ -1,12 +1,12 @@
 import random
 import numpy as np
 
-RANDOM_SEED = random.random()
+RANDOM_SEED = 9 #random.random()
 
 ''' GA parmeters '''
-POPULATION_SIZE = 10
+POPULATION_SIZE = 3
 MUTATION_RATE = 0
-MAX_GENERATIONS = 10
+MAX_GENERATIONS = 15
 CROSSOVER_RATE = 0
 TOURNAMENT_SIZE = 3
 PROBA_SELECTION = 0

### 2021-06-15 run.py

@@ -2,7 +2,7 @@ import matplotlib.pyplot as plt
 import main
 import parameters
 MAX_GENERATIONS = parameters.MAX_GENERATIONS
-
+RANDOM_SEED = parameters.RANDOM_SEED
 # Optional because they are arguments to main
 CROSSOVER_RATE = parameters.CROSSOVER_RATE
 MUTATION_RATE = parameters.MUTATION_RATE 

### 2021-06-15 data.py
enforced Q cst

@@ -22,7 +22,7 @@ them into a coherent dataframe for later statistical and graphical analysis'''
 
 import pandas as pd
 
-def generate_df(generation_history, price_history, mismatch_history, 
+def generate_df(generation_history, price_history, mismatch_history, asset_count_history,
                 dividend_history, random_dividend_history, replacements):
     df = pd.DataFrame()
     
@@ -33,6 +33,7 @@ def generate_df(generation_history, price_history, mismatch_history,
     # Economic variables 
     df["Price"] = price_history
     df["Mismatch"] = mismatch_history
+    df["Q"] = asset_count_history
     df["Div"] = dividend_history
     df["RDiv"] = random_dividend_history
     

### 2021-06-15 main.py

@@ -38,6 +38,9 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     price_history = []
     generation_history = []
     mismatch_history = []
+    asset_count_history = []
+    
+    assetQ = market.count_assets(pop)
     
     # Temp
     agent0_profit = []
@@ -75,7 +78,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         random_dividend_history.append(random_dividend)
         
         ''' B) Apply dividends, interest rate and reinvestment, update profit '''
-        market.wealth_earnings(pop, dividend)
+        # market.wealth_earnings(pop, dividend)
         
         ''' C) Update wealth sum as a function of price '''
         market.update_wealth(pop, price) 
@@ -159,7 +162,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         price_history.append(price)
         
         ''' J) Update inventories ''' 
-        market.update_inventory(pop, price)
+        market.update_inventory(pop, price, assetQ)
         
         ''' K) Record results '''
         maxFitness = max(fitnessValues)
@@ -167,6 +170,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         maxFitnessValues.append(maxFitness)
         meanFitnessValues.append(meanFitness)
         replacements.append(round_replacements)
+        asset_count_history.append(market.count_assets(pop))
         
         print("agg_price before mismatch")
         print(aggregate_ed(price))
@@ -186,7 +190,7 @@ def main(selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         #     for ind in pop:
         #         ind[1] -= 1
 
-        df = data.generate_df(generation_history, price_history, mismatch_history, 
+        df = data.generate_df(generation_history, price_history, mismatch_history, asset_count_history, 
                               dividend_history, random_dividend_history, replacements)
     # return 
     

### 2021-06-15 market.py

@@ -129,28 +129,33 @@ def compute_aggregate_excess_demand(pop, list_excess_demand_func):
         return result
     return aggregate_ed
 
+def count_assets(pop):
+    count = 0
+    for ind in pop:
+        count += ind[3]
+    return count
 
-def update_inventory (pop, price):
+
+def update_inventory (pop, price, assetQ):
     for ind in pop:
         former_asset = ind[3]
         former_loan = ind[4]
         realised_ed = truncate(ind[6] / price - ind[3],4)
-        realised_ed = ind[6] / price - ind[3]
-        print("agent wants up to " + str(realised_ed))
+        print("agent wants up to " + str(realised_ed)) #this is correct
         
         
         ''' If we want to buy assets: non-negative cash buying procedure '''
         cash = ind[2] + ind[3] * price + former_loan - ind[4]
-        ind[3] = 0
         i = 0
         # print("agent wants up to " + str(realised_ed))
         while i < realised_ed:
             if cash - price > 0:
-                ind[3] += 1
-                cash -= price
+                if count_assets(pop) < assetQ:
+                    ind[3] += 1
+                    cash -= price
             i += 1
         ind[2] = cash 
-        # print("agent got " + str(ind[3]))
+        print("agent got " + str(ind[3] - former_asset))
 
         
         # If short selling, set some margin aside

### 2021-06-15 run.py

@@ -15,6 +15,10 @@ print(df)
 # Plot population histograms at the start and at the end
 print("--------------------------")
 
+df.plot(x="Gen", y = ["Price"],
+        kind="line", figsize=(15, 6))
+plt.show()
+
 # print("Initial population was " + str(initial_pop))
 # sns.histplot(data=np.array(initial_pop), legend = False, stat = "density", shrink = 0.85, discrete=True, bins = 11)
 # plt.show()

### 2021-06-28 genetic_algorithm_functions.py
new population.py, types vector

@@ -3,6 +3,7 @@ from deap import creator
 from deap import tools
 from deap import algorithms
 from deap import gp
+import numpy as np
 import random
 from operator import attrgetter
 import parameters
@@ -73,23 +74,31 @@ toolbox.register("generate_vi_individual", tools.initCycle, creator.individual,
 toolbox.register("vi_population_creation", tools.initRepeat, list, toolbox.generate_vi_individual)
 
 def create_mixed_population(POPULATION_SIZE, PROBA_TF, PROBA_VI):
+    global types
+    types = np.array(["NA"])
+    
     def determine_mixed_strategy(PROBA_TF, PROBA_VI):
+        global types
         rd = random.random()
         if rd <= PROBA_TF:
+            types = np.vstack((types, "TF"))
             return toolbox.generate_tf_strategy()
         elif rd > PROBA_TF and rd <= PROBA_TF + PROBA_VI:
+            types = np.vstack((types, "VI"))
             return toolbox.generate_vi_strategy()
     
-    toolbox.register("generate_mix_strategy", determine_mixed_strategy, PROBA_TF, PROBA_VI)
+    toolbox.register("generate_mix_strategy", determine_mixed_strategy, PROBA_TF,
+                     PROBA_VI)
     toolbox.register("generate_mix_individual", tools.initCycle, creator.individual, 
                      (toolbox.generate_mix_strategy, toolbox.generate_wealth, 
                       toolbox.generate_cash, toolbox.generate_asset, 
                       toolbox.generate_loan, toolbox.generate_trading_signal, 
                       toolbox.generate_excess_demand,toolbox.generate_profit,
                       toolbox.generate_ema, toolbox.generate_margin), n=1)
-    toolbox.register("mix_population_creation", tools.initRepeat, list, toolbox.generate_mix_individual)
+    toolbox.register("mix_population_creation", tools.initRepeat, list, 
+                     toolbox.generate_mix_individual)
     pop = toolbox.mix_population_creation(n=POPULATION_SIZE)
-    return pop
+    return pop, types
 
 # Fitness definition
 def ema_fitness(individual):

### 2021-06-28 main.py

@@ -8,6 +8,7 @@ import market_clearing_leap as mc_leap
 import parameters
 import data
 import brownian_motion as bm
+import population_generation as popgen
 
 RANDOM_SEED = parameters.RANDOM_SEED
 random.seed(RANDOM_SEED)
@@ -29,64 +30,12 @@ INITIAL_CASH = parameters.INITIAL_CASH
 INITIAL_ASSETS = parameters.INITIAL_ASSETS
 
 
+
 def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     
-    "we can probably put this as an external function to save space"
-    if mode == "extended":
-        PROBA_TF = parameters.PROBA_TF
-        PROBA_VI = parameters.PROBA_VI
-        PROBA_GP = 0
-    if mode == "open":
-        PROBA_GP = 1
-        PROBA_TF = 0
-        PROBA_VI = 0
-    if mode == "combined":
-        PROBA_GP = parameters.PROBA_GP
-        PROBA_TF = parameters.PROBA_TF
-        PROBA_VI = parameters.PROBA_VI
-        
-### create the populations ###
+    pop_ex, pop_op, balance_sheet, types = popgen.generate_population(mode)
     
-    if PROBA_GP == 1:
-        pop_ex = []
-        # pop_op = gp.create-population(.POPULATIOn_SIZE...)
-    elif PROBA_TF == 1:
-        pop_ex = ga.toolbox.tf_population_creation(n=POPULATION_SIZE)
-        pop_op = []
-    elif PROBA_VI == 1:
-        pop_ex = ga.toolbox.vi_population_creation(n=POPULATION_SIZE)
-        pop_op = []
-    else:
-        if PROBA_GP == 0: 
-            pop_ex = ga.create_mixed_population(POPULATION_SIZE, PROBA_TF, PROBA_VI)
-        else: 
-            # Determine respective population sizes
-            POP_OP_SIZE = 0
-            POP_EX_SIZE = 0
-            for i in range(POPULATION_SIZE):
-                rd = random.random()
-                if rd <= PROBA_GP:
-                    POP_OP_SIZE += 1
-                elif rd > PROBA_GP:
-                    POP_EX_SIZE += 1
-            # Create the two populations
-            pop_ex = ga.create_mixed_population(POP_EX_SIZE, PROBA_TF, PROBA_VI)        
-            # pop_op = gp.create-population(POP_OP_SIZE)
-    pop = pop_ex.copy()
[…]

### 2021-06-28 parameters.py

@@ -11,7 +11,7 @@ PROBA_VI = 0.5
 ''' GA parmeters '''
 POPULATION_SIZE = 4
 MUTATION_RATE = 0.1
-MAX_GENERATIONS = 2
+MAX_GENERATIONS = 3
 CROSSOVER_RATE = 0.1
 TOURNAMENT_SIZE = 3
 PROBA_SELECTION = 0

### 2021-06-28 population_generation.py

@@ -1,0 +1,88 @@
+import parameters
+import numpy as np
+import genetic_algorithm_functions as ga
+import random
+
+INITIAL_CASH = parameters.INITIAL_CASH
+INITIAL_ASSETS = parameters.INITIAL_ASSETS
+PROBA_TF = parameters.PROBA_TF
+PROBA_VI = parameters.PROBA_VI
+PROBA_GP = parameters.PROBA_GP
+POPULATION_SIZE = parameters.POPULATION_SIZE
+
+def generate_population(mode):
+        
+    if mode == "extended":
+        PROBA_TF = parameters.PROBA_TF
+        PROBA_VI = parameters.PROBA_VI
+        PROBA_GP = 0
+    if mode == "open":
+        PROBA_GP = 1
+        PROBA_TF = 0
+        PROBA_VI = 0
+    if mode == "combined":
+        PROBA_GP = parameters.PROBA_GP
+        PROBA_TF = parameters.PROBA_TF
+        PROBA_VI = parameters.PROBA_VI
+        
+### create the populations ###
+
+    types = np.array(["NA"])
+    
+    if PROBA_GP == 1:
+        pop_ex = []
+        # pop_op = gp.create-population(.POPULATIOn_SIZE...)
+        pop_op = [] #temp
+        for i in range(POPULATION_SIZE):
+            types = np.vstack((types, "GP"))
+        
+    elif PROBA_TF == 1:
+        pop_ex = ga.toolbox.tf_population_creation(n=POPULATION_SIZE)
+        pop_op = []
+        for i in range(POPULATION_SIZE):
+            types = np.vstack((types, "TF"))
+    elif PROBA_VI == 1:
+        pop_ex = ga.toolbox.vi_population_creation(n=POPULATION_SIZE)
+        pop_op = []
+        for i in range(POPULATION_SIZE):
+            types = np.vstack((types, "VI"))
+    else:
+        if PROBA_GP == 0: 
+            pop_ex, types = ga.create_mixed_population(POPULATION_SIZE, PROBA_TF, PROBA_VI)
+            pop_op = []
+        else: 
+            # Determine respective population sizes
+            POP_OP_SIZE = 0
+            POP_EX_SIZE = 0
+            for i in range(POPULATION_SIZE):
+                rd = random.random()
+                if rd <= PROBA_GP:
+                    POP_OP_SIZE += 1
+                elif rd > PROBA_GP:
+                    POP_EX_SIZE += 1
[…]

### 2021-06-28 main.py
layout improv

@@ -40,27 +40,12 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     # Create the population and the results accumulators
     generationCounter = 1
     price = INITIAL_PRICE
-    maxFitnessValues = []
-    meanFitnessValues = []
-    replacements = []
-    dividend_history = []
-    random_dividend_history = []
-    price_history = []
     extended_price_history = bm.generate_bm_series(MAX_TIME_HORIZON)
-    plt.plot(extended_price_history)
-    plt.show()
-    generation_history = []
-    mismatch_history = []
-    asset_count_history = []
-    mean_theta = []
-    mean_wealth = []
-    size_pos_pos = []
-    size_neg_pos = []
-    """ we can do a,b,c = [],[],[] to be more concise """
-    
     assetQ = market.count_assets(pop)
-    
     dividend = INITIAL_DIVIDEND
+
+    maxFitnessValues, meanFitnessValues, replacements, dividend_history, random_dividend_history, price_history = [],[],[],[],[],[]   
+    generation_history, mismatch_history, asset_count_history, mean_theta, mean_wealth, size_pos_pos, size_neg_pos = [],[],[],[],[],[], []
     
     print("Initial population")
     print(('{}\n'*len(pop)).format(*pop))
@@ -72,13 +57,7 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     for individual, fitnessValue in zip(pop, fitnessValues):
         individual.fitness.values = fitnessValue
     fitnessValues = [individual.fitness.values[0] for individual in pop]
-    
-    # maxFitness = max(fitnessValues)
-    # meanFitness = sum(fitnessValues) / len(pop)
-    # maxFitnessValues.append(maxFitness)
-    # meanFitnessValues.append(meanFitness)
-
-    
+        
     while generationCounter < MAX_GENERATIONS:
         print("----------------------------------------------------------------------")
         print("Generation " + str(generationCounter))

### 2021-06-28 main.py
Update main.py

@@ -29,8 +29,6 @@ CONSUMPTION_RATE = parameters.CONSUMPTION_RATE
 INITIAL_CASH = parameters.INITIAL_CASH
 INITIAL_ASSETS = parameters.INITIAL_ASSETS
 
-
-
 def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     
     pop_ex, pop_op, balance_sheet, types = popgen.generate_population(mode)

### 2021-06-28 main.py
reworked wealth earnings with new bs

@@ -32,7 +32,8 @@ INITIAL_ASSETS = parameters.INITIAL_ASSETS
 def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     
     pop_ex, pop_op, balance_sheet, types = popgen.generate_population(mode)
-    
+    print("initialised bs")
+    print(balance_sheet)
     pop = pop_ex.copy() #temp
     
     # Create the population and the results accumulators
@@ -51,10 +52,11 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     print("[Theta Wealth Cash Asset Loan TradingSignal RawExcessDemand     Profit     EMA profit   Margin]")
     print("[ 0       1     2    3     4         5             6           7            8              9]")
     
-    fitnessValues = list(map(ga.toolbox.evaluate, pop))
-    for individual, fitnessValue in zip(pop, fitnessValues):
-        individual.fitness.values = fitnessValue
-    fitnessValues = [individual.fitness.values[0] for individual in pop]
+    # "Put this as a single function"
+    # fitnessValues = list(map(ga.toolbox.evaluate, pop))
+    # for individual, fitnessValue in zip(pop, fitnessValues):
+    #     individual.fitness.values = fitnessValue
+    # fitnessValues = [individual.fitness.values[0] for individual in pop]
         
     while generationCounter < MAX_GENERATIONS:
         print("----------------------------------------------------------------------")
@@ -68,14 +70,20 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         print("dividend is " + str(dividend))
         random_dividend_history.append(random_dividend)
         
+        print("bs before B")
+        print(balance_sheet)
         ''' B) Apply dividends, interest rate and reinvestment, update profit '''
         market.wealth_earnings(pop, dividend)
+        market.bs_wealth_earnings(balance_sheet, dividend)
+        print("bs after B")
+        print(balance_sheet)
+        
+        print(('{}\n'*len(pop)).format(*pop))
         
         ''' C) Update wealth and margin as a function of price '''
         market.consumption(pop, CONSUMPTION_RATE, price)
         market.update_wealth(pop, price) 
         market.update_margin(pop, price)
-    
         
         ''' D) Hypermutation operator '''
         pop, round_replacements = ga.hypermutate(pop)
@@ -83,11 +91,11 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
[…]

### 2021-06-28 market.py

@@ -59,8 +59,8 @@ def draw_dividend(DIVIDEND_GROWTH_RATE):
     dividend = truncate(dividend, 3)
     return dividend, random_dividend
 
-        
-def wealth_earnings(pop, dividend):
+"""TO REMOVE"""
+def wealth_earnings(pop, dividend): 
     # print(INTEREST_RATE)
     # print(REINVESTMENT_RATE)
     for ind in pop:
@@ -72,6 +72,14 @@ def wealth_earnings(pop, dividend):
         ind[2] += REINVESTMENT_RATE * (INTEREST_RATE * ind[2] + dividend * ind[3])
         ind[2] = truncate(ind[2],3)
     return ind
+
+def bs_wealth_earnings(balance_sheet, dividend):
+    for row in balance_sheet:
+        # Update Profit
+        row[6] = truncate(REINVESTMENT_RATE * (INTEREST_RATE * row[1] + dividend * row[2]),3)
+        # Update Cash
+        row[1] += REINVESTMENT_RATE * (INTEREST_RATE * row[1] + dividend * row[2])
+        row[1] = truncate(row[1],3)
 
 def update_wealth(pop, price):
     for ind in pop:

### 2021-06-29 market_clearing_leap.py
testing leap

@@ -108,4 +108,11 @@ def solver_linear_shortcut(pop, price):
         
     return new_price 
 
-
+""" Testing 
+def f(x):
+    return sum(x)**2
+best_genome = ea_solve_noverbose(f,
+          bounds=[(1.1111, 10)], generations = 50, pop_size = 500,
+          mutation_std=0.1, hard_bounds = True) #max = False
+print(best_genome)
+"""

### 2021-06-29 draft.tex
Updates from Overleaf

@@
 \DeclareMathOperator*{\argmin}{arg\,min}
-\usepackage{algorithm}% http://ctan.org/pkg/algorithms 
-\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx 
-\newcommand{\var}[1]{{\ttfamily#1}}% variable
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\newcommand{\var}[1]{{\ttfamily#1}}
+\newcommand\boldred[1]{\textcolor{red}{\textbf{#1}}}
 
-\title{Evolutionary market ecologies \\ Model \& Numerical experiments}
+\title{Evolutionary market ecologies and market malfunction\\
+Draft}
 \author{Aymeric Vi\'{e}}
@@
 
-\section{Trend following Model description}
+\begin{abstract}
+    \boldred{Abstract TBA}
+\end{abstract}
+
+\section{Introduction}
+\boldred{Introduction TBA}
+\boldred{Everything related to GP and multi asset: in footnotes}
+% Maybe paragraphs could be
+% market ecology
+% evolutionary computation
+% evolution in financial markets
+
+\section{Model}
 
 \subsection{Initialisation}
+\boldred{To update with new structure}
 
@@
 The market is initialised with an initial price of $100$ that matches the fundamental value of the asset, and initial dividends of $100$. 
+\boldred{Add a note for the multi asset thing.
+We actually need to think about multiple assets now that we are doing these big changes.
+Notably, the ED and TS matrix would change to be $n\timesK$ matrices, where n pop size and K number of assets.
+The agent strategies change as well to be a K-vector of strategies.
+It would be better to have a very general case, a very general structure, and so that we have the right structure and only need particular settings or specific small fixes to make it work and extend it.}
 
 \subsection{Strategy and market behavior}
+\boldred{Update for the other strategies as well}
 At each time period, trend followers look at past prices to decide the content of their next portfolio, choosing between cash and an asset.
@@
 
+\boldred{With multiple assets, how do we formulate the excess demand? Replace E and TS and p and S by the matrix form (generalised ED function)?}
+
 \begin{equation}
@@
 \end{equation}
+\boldred{Extends to a vector S p with multiple assets}
 
 \subsection{Market clearing}
+\boldred{Add something to imagine clearing with a multi asset environment, just consists in adding all the assets with each its own price}
 
@@
 \end{equation}
+
+We impose a non-negativity constraint on the cash. 
+% In other words, the agents cannot enter into cash debt to buy more assets. 
+This must work with leverage.
 
@@
 
-The non-lerning case (replacement dynamics only) can be illustrated with:
+The non-learning case (replacement dynamics only) can be illustrated with:
 \begin{itemize}

### 2021-06-29 testing_&_archiv/temp_solver.py
Create temp_solver.py

@@ -1,0 +1,76 @@
+import numpy as np
+
+'''
+Here is a simple optimisation solver for the market clearing algorithm. 
+We minimise the squared of aggregate excess demand, under the constraint of
+a Limit-Up-Limit-Down (LU-LD) circuit breaker, which limits prices to 
+[LD*p(t-1), LU*p(t-1)], with LD=1/2 and LU = 2
+
+
+The solver uses the LEAP package: https://github.com/AureumChaos/LEAP
+It is a package focusing on evolutionary algorithms, that has an amazing 
+ea_solve function that ... solves a real valued function.
+'''
+
+from leap_ec import Individual, Representation
+from leap_ec import ops, probe
+from leap_ec.algorithm import generational_ea
+from leap_ec.problem import FunctionProblem
+from leap_ec.real_rep import create_real_vector
+from leap_ec.real_rep.ops import mutate_gaussian
+
+"""Provides a simple, top-level interfact that optimizes a real-valued
+function using a simple generational EA.
+:param function: the function to optimize; should take lists of real
+    numbers as input and return a float fitness value
+:param [(float, float)] bounds: a list of (min, max) bounds to define the
+    search space
+:param int generations: the number of generations to run for
+:param int pop_size: the population size
+:param float mutation_std: the width of the mutation distribution
+:param bool maximize: whether to maximize the function (else minimize)
+:param bool viz: whether to display a live best-of-generation plot
+:param bool hard_bounds: if True, bounds are enforced at all times during
+evolution; otherwise they are only used to initialize the population. """
+
+def ea_solve_noverbose(function, bounds, generations=100, pop_size=2,
+             mutation_std=1.0, maximize=False, viz=False, viz_ylim=(0, 1),
+             hard_bounds=True):
+
+    if hard_bounds:
+        mutation_op = mutate_gaussian(std=mutation_std, hard_bounds=bounds,
+                        expected_num_mutations='isotropic')
+    else:
+        mutation_op = mutate_gaussian(std=mutation_std,
+                        expected_num_mutations='isotropic')
+    pipeline = [
+        ops.tournament_selection,
+        ops.clone,
+        mutation_op,
+        ops.uniform_crossover(p_swap=0.4),
+        ops.evaluate,
+        ops.pool(size=pop_size)
+    ]
+    ea = generational_ea(max_generations=generations,
+                         pop_size=pop_size,
+                         problem=FunctionProblem(function, maximize),
[…]

### 2021-06-30 main.py
reworked C wealth update, consumption, margin update

@@ -39,7 +39,7 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     # Create the population and the results accumulators
     generationCounter = 1
     price = INITIAL_PRICE
-    extended_price_history = bm.generate_bm_series(MAX_TIME_HORIZON)
+    extended_price_history = bm.generate_bm_series(MAX_TIME_HORIZON+1)
     assetQ = market.count_assets(pop)
     dividend = INITIAL_DIVIDEND
 
@@ -70,20 +70,17 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         print("dividend is " + str(dividend))
         random_dividend_history.append(random_dividend)
         
-        print("bs before B")
-        print(balance_sheet)
+        # print("bs before B")
+        # print(balance_sheet)
         ''' B) Apply dividends, interest rate and reinvestment, update profit '''
-        market.wealth_earnings(pop, dividend)
+        market.wealth_earnings(pop, dividend) #temp
         market.bs_wealth_earnings(balance_sheet, dividend)
-        print("bs after B")
-        print(balance_sheet)
-        
-        print(('{}\n'*len(pop)).format(*pop))
         
         ''' C) Update wealth and margin as a function of price '''
-        market.consumption(pop, CONSUMPTION_RATE, price)
-        market.update_wealth(pop, price) 
-        market.update_margin(pop, price)
+        market.consumption(pop, CONSUMPTION_RATE, price) #temp
+        market.update_wealth(pop, price)  #temp
+        market.update_margin(pop, price) #temp
+        market.bs_wealth_update(balance_sheet, price, CONSUMPTION_RATE)
         
         ''' D) Hypermutation operator '''
         pop, round_replacements = ga.hypermutate(pop)

### 2021-06-30 market.py

@@ -81,11 +81,14 @@ def bs_wealth_earnings(balance_sheet, dividend):
         row[1] += REINVESTMENT_RATE * (INTEREST_RATE * row[1] + dividend * row[2])
         row[1] = truncate(row[1],3)
 
+
+""" TO REMOVE """
 def update_wealth(pop, price):
     for ind in pop:
         ind[1] = truncate(ind[2] + ind[3] * price  - ind[4],3)
     return ind
 
+""" TO REMOVE """
 def consumption(pop, CONSUMPTION_RATE, price):
     for ind in pop:
         if ind[2] >= CONSUMPTION_RATE:
@@ -102,7 +105,54 @@ def consumption(pop, CONSUMPTION_RATE, price):
                     # sells
                 if ind[3] * price < consume_amount:
                     ind[1] -= 1_000
-            
+
+""" TO REMOVE """
+def update_margin(pop, price):
+    for ind in pop:
+        if ind[3] < 0:
+            ind[2] += ind[9]
+            ind[9] = 0
+            ind[9] = abs(ind[3]) * price
+            ind[2] -= ind[9]
+            if ind[2] < 0:
+                ''' If not enough cash for margin, and since we are in debt, 
+                we are automatically replaced '''
+                ind[1] = -1_000
+    return ind
+
+def bs_wealth_update(balance_sheet, price, CONSUMPTION_RATE):
+    for row in balance_sheet:
+        
+        # Apply consumption
+        if row[1] >= CONSUMPTION_RATE:
+            row[1] -= CONSUMPTION_RATE
+        if row[1] < CONSUMPTION_RATE:
+            row[1] = 0
+            consume_amount = CONSUMPTION_RATE - row[1]
+            if row[2] <= 0:
+                row[1] -= 1_000
+            if row[2] > 0:
+                # agent will try to sell assets to fund consumption
+                if row[2] * price >= consume_amount:
+                    row[2] -= consume_amount / price
+                    # sells
+                if row[2] * price < consume_amount:
+                    row[1] -= 1_000       
+
+        # Update margin
+        if row[2] < 0:
+            row[1] += row[8]
+            row[8] = 0
+            row[8] = abs(row[2]) * price
+            row[1] -= row[8]
+            if row[1] < 0:
+                ''' If not enough cash for margin, and since we are in debt, 
+                we are automatically replaced '''
+                row[1] = -1_000
+                
+        # Update wealth
[…]

### 2021-06-30 genetic_algorithm_functions.py
wip hymerputation reworked

@@ -26,6 +26,8 @@ INITIAL_ASSETS = parameters.INITIAL_ASSETS
 INITIAL_CASH = parameters.INITIAL_CASH
 MIN_VALUATION = parameters.MIN_VALUATION
 MAX_VALUATION = parameters.MAX_VALUATION
+PROBA_TF = parameters.PROBA_TF
+PROBA_VI = parameters.PROBA_VI
 
 
 toolbox = base.Toolbox()
@@ -56,7 +58,7 @@ toolbox.register("tf_population_creation", tools.initRepeat, list, toolbox.gener
 
 toolbox.register("generate_no_asset", random.randint, 0, 0)
 
-toolbox.register("generate_hyper_individual", tools.initCycle, creator.individual, 
+toolbox.register("generate_hyper_tf_individual", tools.initCycle, creator.individual, 
                  (toolbox.generate_tf_strategy, toolbox.generate_wealth, 
                   toolbox.generate_cash, toolbox.generate_no_asset, 
                   toolbox.generate_loan, toolbox.generate_trading_signal, 
@@ -72,6 +74,26 @@ toolbox.register("generate_vi_individual", tools.initCycle, creator.individual,
                   toolbox.generate_excess_demand,toolbox.generate_profit,
                   toolbox.generate_ema, toolbox.generate_margin), n=1)
 toolbox.register("vi_population_creation", tools.initRepeat, list, toolbox.generate_vi_individual)
+
+toolbox.register("generate_hyper_vi_individual", tools.initCycle, creator.individual, 
+                 (toolbox.generate_vi_strategy, toolbox.generate_wealth, 
+                  toolbox.generate_cash, toolbox.generate_no_asset, 
+                  toolbox.generate_loan, toolbox.generate_trading_signal, 
+                  toolbox.generate_excess_demand,toolbox.generate_profit,
+                  toolbox.generate_ema, toolbox.generate_margin), n=1)
+
+def determine_mixed_strategy(PROBA_TF, PROBA_VI):
+    global types
+    rd = random.random()
+    if rd <= PROBA_TF:
+        types = np.vstack((types, "TF"))
+        return toolbox.generate_tf_strategy()
+    elif rd > PROBA_TF and rd <= PROBA_TF + PROBA_VI:
+        types = np.vstack((types, "VI"))
+        return toolbox.generate_vi_strategy()
+
[…]

### 2021-06-30 main.py

@@ -85,6 +85,8 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         ''' D) Hypermutation operator '''
         pop, round_replacements = ga.hypermutate(pop)
         ga.fitness_for_invalid(pop)
+        print(('{}\n'*len(pop_ex)).format(*pop_ex))
+        print(type(pop_ex))
         
         ''' E) Deduce fitness as EMA ''' 
         market.compute_ema(pop)
@@ -190,6 +192,6 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     # return 
     # return price, initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df
     print(('{}\n'*len(pop)).format(*pop))
-    return df, extended_price_history
+    return df, extended_price_history, pop_ex
 
 

### 2021-06-30 run.py

@@ -12,7 +12,7 @@ MUTATION_RATE = parameters.MUTATION_RATE
 PROBA_SELECTION = parameters.PROBA_SELECTION
 
 # price, initial_pop, pop, maxFitnessValues, meanFitnessValues, replacements, agent0_profit, agent0_ema, dividend_history, price_history, random_dividend_history, list_excess_demand_func, aggregate_ed, df = main.main(PROBA_SELECTION, CROSSOVER_RATE, MUTATION_RATE)
-df, extended_price_history = main.main("extended", PROBA_SELECTION, CROSSOVER_RATE, MUTATION_RATE)
+df, extended_price_history, pop_ex = main.main("extended", PROBA_SELECTION, CROSSOVER_RATE, MUTATION_RATE)
 
 print(df)
 print("--------------------------")
@@ -34,3 +34,10 @@ plt.show()
 
 
 print("-----------------------")
+
+
+print(('{}\n'*len(pop_ex)).format(*pop_ex))
+print(type(pop_ex))
+print(type(pop_ex[0]))
+pop = pop_ex.copy()
+

### 2021-06-30 run.py
tests for pop objects

@@ -39,5 +39,18 @@ print("-----------------------")
 print(('{}\n'*len(pop_ex)).format(*pop_ex))
 print(type(pop_ex))
 print(type(pop_ex[0]))
-pop = pop_ex.copy()
 
+pop = pop_ex.copy()
+del pop[0]
+print(('{}\n'*len(pop)).format(*pop))
+print("Deleting is succesful")
+
+import genetic_algorithm_functions as ga
+pop = pop_ex.copy()
+del pop[0]
+add = ga.toolbox.generate_hyper_tf_individual()
+print(add)
+pop.append(add)
+print(('{}\n'*len(pop)).format(*pop))
+print(type(pop_ex[3]))
+print("Deleting and adding a new DEAP individual is succesful")

### 2021-07-01 genetic_algorithm_functions.py
finish hyper2.0 and bug with NaN extended price history for log in TS update

@@ -7,6 +7,7 @@ import numpy as np
 import random
 from operator import attrgetter
 import parameters
+import sys
 
 
 # # =============================================================================
@@ -214,7 +215,6 @@ def hypermutate(pop):
 toolbox.register("hypermutate", hypermutate)
 
 # Define the hypermutation (insolvency) parameter
-# round_replacements = 0
 def hypermutate2(pop_ex, pop_op, types, balance_sheet, mode):
     
     if mode == "extended":
@@ -229,13 +229,6 @@ def hypermutate2(pop_ex, pop_op, types, balance_sheet, mode):
         PROBA_GP = parameters.PROBA_GP
         PROBA_TF = parameters.PROBA_TF
         PROBA_VI = parameters.PROBA_VI
-        
-    pop_ex_temp = list(map(toolbox.clone, pop_ex))
-    pop_op_temp = list(map(toolbox.clone, pop_op))
-    types_temp = types.copy()
-    balance_sheet_temp = balance_sheet.copy()
-    
-    """ FFS POP SIZES BTW OP AND EX MAY CHANGE HERE """
     
     round_replacements = 0
     
@@ -244,64 +237,69 @@ def hypermutate2(pop_ex, pop_op, types, balance_sheet, mode):
         if balance_sheet[i][0] <= 0:
             
             """ Agent is insolvent. Delete """
-            del pop_ex[i].fitness.values
-            del pop_ex[i]
+            if i < len(pop_ex):
+                del pop_ex[i].fitness.values
+                del pop_ex[i]
+            if i >= len(pop_ex):
+                del pop_op[i - len(pop_ex)].fitness.values
+                del pop_op[i - len(pop_ex)]
+                
             balance_sheet = np.delete(balance_sheet, (i), axis=0)
             types = np.delete(types, (i), axis=0)
-            
             
             """ Replace """
             ind_bs = np.array([0, INITIAL_CASH, INITIAL_ASSETS, 0, 0, 0, 0, 0, 0])
             balance_sheet = np.vstack((balance_sheet, ind_bs))
             
             # Draw the type to choose what pop and type to append
-            """ Will deap accept to append something to the pop??"""
             if PROBA_GP == 1:
                 """ The incoming agent is of type GP """
                 # add = gp.create-population(.POPULATIOn_SIZE...)
                 # pop_op.append(add)
                 types = np.vstack((types, "GP"))
+                round_replacements += 1
                 
             elif PROBA_TF == 1:
[…]

### 2021-07-01 main.py

@@ -40,6 +40,7 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
     generationCounter = 1
     price = INITIAL_PRICE
     extended_price_history = bm.generate_bm_series(MAX_TIME_HORIZON+1)
+    extended_price_history = [abs(x) for x in extended_price_history]
     assetQ = market.count_assets(pop)
     dividend = INITIAL_DIVIDEND
 
@@ -85,8 +86,14 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         ''' D) Hypermutation operator '''
         pop, round_replacements = ga.hypermutate(pop)
         ga.fitness_for_invalid(pop)
+        print("before hyp")
         print(('{}\n'*len(pop_ex)).format(*pop_ex))
         print(type(pop_ex))
+        pop_ex, pop_op, types, balance_sheet, round_replacements = ga.hypermutate2(pop_ex, pop_op, types, balance_sheet, mode)
+        print("after hyp")
+        print(('{}\n'*len(pop_ex)).format(*pop_ex))
+        print(types)
+        print(balance_sheet)
         
         ''' E) Deduce fitness as EMA ''' 
         market.compute_ema(pop)

### 2021-07-01 market.py

@@ -171,6 +171,9 @@ def update_trading_signal(pop, extended_price_history):
     for ind in pop:
         if len(extended_price_history) > 1:
             if len(extended_price_history) > ind[0]:
+                print("Extended Price History")
+                print(extended_price_history)
+                print(np.log(extended_price_history))
                 ind[5] = truncate(np.log2(extended_price_history[-1]) - np.log2(extended_price_history[-ind[0]]),3)
             if len(extended_price_history) <= ind[0]:
                 # The trader does not have the information to run her strategy. She waits in indifference.

### 2021-07-01 run.py

@@ -36,21 +36,3 @@ plt.show()
 print("-----------------------")
 
 
-print(('{}\n'*len(pop_ex)).format(*pop_ex))
-print(type(pop_ex))
-print(type(pop_ex[0]))
-
-pop = pop_ex.copy()
-del pop[0]
-print(('{}\n'*len(pop)).format(*pop))
-print("Deleting is succesful")
-
-import genetic_algorithm_functions as ga
-pop = pop_ex.copy()
-del pop[0]
-add = ga.toolbox.generate_hyper_tf_individual()
-print(add)
-pop.append(add)
-print(('{}\n'*len(pop)).format(*pop))
-print(type(pop_ex[3]))
-print("Deleting and adding a new DEAP individual is succesful")

### 2021-07-01 main.py
hypermut2.0 corrected

@@ -84,6 +84,9 @@ def main(mode, selection_proba, CROSSOVER_RATE, MUTATION_RATE):
         market.bs_wealth_update(balance_sheet, price, CONSUMPTION_RATE)
         
         ''' D) Hypermutation operator '''
+        
+        balance_sheet[0,0] = 0
+        
         pop, round_replacements = ga.hypermutate(pop)
         ga.fitness_for_invalid(pop)
         print("before hyp")

### 2021-07-01 market.py

@@ -171,9 +171,9 @@ def update_trading_signal(pop, extended_price_history):
     for ind in pop:
         if len(extended_price_history) > 1:
             if len(extended_price_history) > ind[0]:
-                print("Extended Price History")
-                print(extended_price_history)
-                print(np.log(extended_price_history))
+                # print("Extended Price History")
+                # print(extended_price_history)
+                # print(np.log(extended_price_history))
                 ind[5] = truncate(np.log2(extended_price_history[-1]) - np.log2(extended_price_history[-ind[0]]),3)
             if len(extended_price_history) <= ind[0]:
                 # The trader does not have the information to run her strategy. She waits in indifference.

### 2021-07-05 draft.tex

@@
 \begin{abstract}
-    \boldred{Abstract TBA}
+    In financial markets, distributions of strategies influence market malfunction: misspricing, volatility and inefficiency.
+Modelling the economy as an ecology of strategies gives insights into market behavior, returns and long term evolution.
+    Economic agents are continuously readjusting to the dynamics their interactions create, exploring a large space of possible actions.
+    How does this adaptation of trading strategies influence the dynamics of markets and strategies? Does learning improve or undermine a market path to efficiency and equilibrium?
+    Here, we study financial adaptation as the evolution of trading strategies using evolutionary algorithms methods in an artificial stock market, and its impact on market and ecology outcomes. 
+    \footnote{\textcolor{gray}{Insert summary of main results, contribution of main results to previous knowledge, and more general context of results}}
 \end{abstract}
@@
 \section{Introduction}
-\boldred{Introduction TBA}
+
+\subsection{Market ecology and malfunction/dynamcs}
+\subsection{Learning in financial markets}
+\subsection{Research questions and motivation}
+\subsection{Model}
+\subsection{Main results and implications}
+
+\boldred{Introduction TBD}
 \boldred{Everything related to GP and multi asset: in footnotes}
@@
 We actually need to think about multiple assets now that we are doing these big changes.
-Notably, the ED and TS matrix would change to be $n\timesK$ matrices, where n pop size and K number of assets.
+Notably, the ED and TS matrix would change to be $nx$ matrices, where n pop size and K number of assets.
 The agent strategies change as well to be a K-vector of strategies.

### 2021-07-06 draft.tex
Updates from Overleaf

@@
 
-\subsection{Market ecology and malfunction/dynamcs}
+\subsection{Market ecology and malfunction/dynamics}
+\boldred{related work and gap that we fill}
 \subsection{Learning in financial markets}
+\boldred{related work and gap that we fill}
 \subsection{Research questions and motivation}
@@
 \subsection{Main results and implications}
+What he hope to show:
+\begin{itemize}
+    \item Taking learning into account changes the stability regions of the model (regions of the strategy space that we thought to be stable/unstable are actually different)
+    \item Taking learning into accounts allows to get more realistic in terms of stylized facts
+    \item Learning shows more complex behaviors, cycles and complexity
+    \item \boldred{policy implications}
+\end{itemize}
 
@@
 \boldred{Everything related to GP and multi asset: in footnotes}
+\boldred{a literature review}
 % Maybe paragraphs could be
